---
title: "Supplementary Information"
output: pdf_document
header-includes: 
  \usepackage{booktabs}
  \usepackage{microtype}
  \usepackage{longtable}
  \usepackage[margin=1in]{geometry}
---

## Evolutionary simulations of _Z_-linked suppression gene drives

## Supplementary Figures and Supplementary Tables

All of the figures and tables in this document can also be viewed online at **XXXXX**, along with the R code used to generate them.

\newpage


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)
library(kableExtra)
library(stringr)

make_table <- function(file, LT = TRUE, digits = getOption("digits")){
  kable(readRDS(file), "latex", longtable = LT, booktabs = TRUE, digits = digits) %>%
    kable_styling(latex_options = c("repeat_header", "scale_down"), font_size = 7)
}
```

# Supplementary figures


<!-- \begin{figure} -->
<!-- \centering -->
<!-- \includegraphics{../figures/figureS1.pdf} -->
<!-- \end{figure} -->
<!-- **Figure S1**: \footnotesize{The percentage of simulations of a female-sterilising $Z^*$ allele (not a \textit{W}-shredder) that ended in extinction, for all runs with a particular value (shown on the \textit{x}-axis) for a given parameter (shown in the panels). Note that the maximum values on the $y$-axis are more smaller than in Figure 2, indicating that a female-sterilising $Z^*$ allele is likely to be less effective than a \textit{W}-shredder at suppressing populations.  The panels are ordered by the range of the \textit{x}-axis, which indicates the relative importance of the variables for the probability of extinction.} -->


```{r fig.width = 12, fig.height = 14}
readRDS("../figures/S1_fig.rds")
```
**Figure S1**: The percentage of simulations of a female-sterilising $Z^*$ allele (not a \textit{W}-shredder) that ended in extinction, for all runs with a particular value (shown on the \textit{x}-axis) for a given parameter (shown in the panels). Note that the maximum values on the $y$-axis are more smaller than in Figure 2, indicating that a female-sterilising $Z^*$ allele is likely to be less effective than a \textit{W}-shredder at suppressing populations.  The panels are ordered by the range of the \textit{x}-axis, which indicates the relative importance of the variables for the probability of extinction.
\newpage

```{r fig.width = 12, fig.height = 14}
readRDS("../figures/S2_fig.rds")
```
**Figure S2**: The percentage of simulations of a female-sterilising $Z^*$ allele (not a \textit{W}-shredder) that ended in extinction, for all runs with a particular value (shown on the \textit{x}-axis) for a given parameter (shown in the panels). Note that the maximum values on the $y$-axis are more smaller than in Figure 2, indicating that a female-sterilising $Z^*$ allele is likely to be less effective than a \textit{W}-shredder at suppressing populations.  The panels are ordered by the range of the \textit{x}-axis, which indicates the relative importance of the variables for the probability of extinction.
\newpage

```{r fig.width = 12, fig.height = 14}
readRDS("../figures/S3_fig.rds")
```
**Figure S3**: The percentage of simulations of a female-sterilising $Z^*$ allele (not a \textit{W}-shredder) that ended in extinction, for all runs with a particular value (shown on the \textit{x}-axis) for a given parameter (shown in the panels). Note that the maximum values on the $y$-axis are more smaller than in Figure 2, indicating that a female-sterilising $Z^*$ allele is likely to be less effective than a \textit{W}-shredder at suppressing populations.  The panels are ordered by the range of the \textit{x}-axis, which indicates the relative importance of the variables for the probability of extinction.





\newpage
# Supplementary tables

```{r echo=FALSE}
count_S1 <- sum(readRDS("../figures/tableS1.rds")[,2])
count_S2 <- sum(readRDS("../figures/tableS2.rds")[,2])
```


**Table S1**: The number and percentage of simulation runs (out of `r count_S1` total) that ended with the five possible outcomes, for the subset of simulation runs focusing on a _W_-shredder gene drive.
```{r, results='asis'}
make_table("../figures/tableS1.rds")
```
\newpage

**Table S2**: The number and percentage of simulation runs (out of `r count_S2` total) that ended with the five possible outcomes, for the subset of simulation runs focusing on a female-sterilising _Z_-linked gene drive.
```{r, results='asis'}
make_table("../figures/tableS2.rds")
```
