---
title: Evolutionary simulations of _Z_-linked suppression gene drives
author:
  - name: Luke Holman
    affiliation: "1"
address:
  - code: "1"
    address: School of BioSciences, The University of Melbourne, Victoria 3010, Australia.

corresp_author_name:  "L. Holman"
corresp_author_email: "luke.holman@unimelb.edu.au"

subject:
  - "Evolutionary biology"
  - "Theoretical modelling"
  - "Gene drives"

keywords:
  - Sex chromosomes
  - Gene drives
  - Population control
  - Schistosomiasis
  - Selfish genes

abstract: |
  Synthetic gene drives may soon be used to suppress or eliminate populations of disease vectors, pathogens, invasive species, and agricultural pests. Recent proposals have focused on using _Z_-linked gene drives to control species with _ZW_ sex determination, which include Lepidopteran pests and parasitic trematodes. These proposals include _Z_-linked '_W_-shredders', which would suppress populations by cleaving the _W_ chromosome and causing females to produce only sons, as well as _Z_-linked female-sterilising gene drives. Here I use eco-evolutionary simulations to evaluate the potential of some proposed _Z_-linked gene drives, and to produce recommendations regarding their design and use. The simulations show that _W_-shredders are likely to be highly effective at eradicating populations provided that resistance to _W_-shredding cannot evolve. However, _W_-shredder alleles can invade populations from very low frequencies, making it difficult to suppress specific populations while leaving others untouched; this issue may restrict their possible uses.
  
author_contributions: |
  LH performed the analyses and wrote the manuscript.
  
data_accessibility: |
  A website presenting all R scripts used to the run the simulation and analyse the data can be found at https://lukeholman.github.io/W_shredder/. 

conflict_of_interest: |
  The author declares no conflict of interest.

funding: |
  This project was stimulated by an ESEB _Progress Meetings in Evolutionary Biology_ meeting, funded by grants from ESEB (European Society for Evolutionary Biology) and from the Swiss National Science Foundation.

acknowledgements: |
  I thank the organisers (Anna Lindholm and Tom Price), funding bodies (European Society for Evolutionary Biology; Swiss National Science Foundation), and attendees of the 2018 ESEB _Progress Meetings in Evolutionary Biology_, which provided the impetus for this paper. I also thank Kevin Esvelt and colleagues for describing their ongoing research on a personal webpage; their ideas were instrumental to this paper.

bibliography: references.bib

## change to true to add optional line numbering
lineno: true

site: bookdown::bookdown_site
output: 
  bookdown::pdf_book:
    base_format: rticles::rsos_article
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
source("../code/table_1_code.r")
```

# Introduction

Developments in genetic engineering will soon make it feasible to alter or eliminate populations of disease vectors, pathogens, agricultural pests, and invasive species using 'gene drives' [@gantz2015hi; @hammond2016cr; @wang2016cr; @prowse2017do; @kyrou2018cr; @noble2018cu]. Gene drives cause particular alleles (usually transgenes) to propagate through populations via a range of mechanisms, which include gene conversion, poison-antidote systems, segregation distortion, and genetic incompatibility [@lindholm2016ec; @champer2016ch; @oberhofer2019cl]. For example, CRISPR-Cas9 gene editing can be used to create a transgenic insertion that is transmitted to almost 100% of the offspring of heterozygous individuals instead of the usual 50%; this type of gene drive functions by inducing a double-stranded DNA break at the homologous wild type locus, which is then repaired using the transgene as a template. Gene drives are often categorised into two types: replacement drives, which aim to spread a human-beneficial allele throughout a population (e.g. a mosquito allele that interferes with the tranmission of malaria [@gantz2015hi; @marshall2015gene]), and suppression drives, which reduce the size of a population (potentially to extinction). Suppression drives typically work by using non-Mendelian inheritance to spread alleles that cause lethality or sterility [@hammond2016cr; @kyrou2018cr; @maselko2018ge], or skew the offspring sex ratio -- typically towards males [@windbichler2008ta; @galizi2014sy; @beaghton2017ve; @burt2018se; @papathanos2018re]. 

Recent theoretical papers have investigated the feasibility, efficacy, and potential negative consequencies of various types of gene drives. For example, Noble et al. [@noble2018cu] showed that the basic version of a CRISPR-Cas9 gene drive might be highly invasive and could rapidly spread to fixation across whole species, which is often an undesirable outcome. Conversely, other models have concluded that gene drives are likely to fail if populations can evolve resistance to their effects [@drury2017cr; @unckless2017ev]. The issue of resistance is compounded because the standard implementation of CRISPR-Cas9 gene drive (but perhaps not updated versions; [@esvelt2014em; @unckless2017ev; @prowse2017do; @kyrou2018cr]) tends to create its own resistance alleles, e.g. when the double-stranded break induced by Cas9 is repaired using an alternative DNA repair pathway (non-homologous end joining; NHEJ) instead of homology-directed repair [@gantz2015mu; @gantz2015hi; @hammond2016cr; @wang2016cr; @unckless2017ev]. Given the potential safety, ethical, and sociopolitical concerns surrounding gene drives, some models have focused on gene drives that would go extinct after a time [@min2017da; @burt2018se; @noble2019da], would stay confined to particular populations [@maselko2018ge; @noble2019da], and/or could be reversed once they have spread [@vella2017ev]. 

Here, I focus on the evolutionary dynamics of _Z_-linked suppression gene drives. The simulation is inspired by proposals for various types of _Z_-linked gene drives by Kevin Esvelt and colleagues, as well as ongoing efforts to develop these _Z_ drives (see [www.sculptingevolution.org](); at the time of writing, these ideas have not been published elsewhere). Various _Z_-linked suppression drives proposed by Esvelt and colleagues are shown schematically in Figure 1. Depending on its design, mode of action and the biology of the target species, _Z_ chromosomes carrying the drive allele (denoted _Z\*_) might enjoy a transmission advantage in _Z\*W_ females (Figure 1B, and perhaps also 1C), and optionally also in _Z\*Z_ males. Esvelt et al. focus on using _Z_ drives to control the _Schistosoma_ trematodes responsible for schistosomiasis, though _Z_ drives could theoretically be used to control any organism with female-heterogametic sex determination, such as Lepidopteran agricultural pests, invasive populations of cane toads _Bufo marinus_ [@abramyan2009z], or even invasive birds.

A _Z_-linked gene drive could suppress populations by biasing gametogenesis in females, for example by inducing double-stranded DNA breaks in the _W_ chromosome in order to inactivate it; such a gene drive would be a '_W_-shredder', analagous to the _X_- and _Y_-shredders under development for _XY_ species [@windbichler2008ta; @north2013mo; @galizi2014sy; @burt2018se; @papathanos2018re; @prowse2019]. Females carrying the gene drive would thus produce relatively few viable _W_-bearing eggs, and therefore produce mainly drive-carrying sons. Esvelt et al. point out that the evolutionary dynamics of the drive will depend on the fitness of drive carriers relative to wild types, the timing of _W_-shredding (e.g. in germ cells, ova, or zygotes), and the ecology of the target species. For example, some _W_-shredder designs might allow drive females to produce roughly the same number of (mostly-male) offspring as a wild-type female provided that the _W_ chromosome is destroyed early enough in oogenesis/development that the lost daughters can be replaced by sons (Figure 1B). Alternatively, drive-carrying females might produce half the number of offspring (or less), e.g. if the drive works by destroying all ova or offspring that carry a _W_ chromosome, and this loss is not compensated by reduced competition on the surviving offspring. Esvelt et al. also proposed that one could suppress populations using a _Z_-linked locus that caused sterility or lethality in females, either by shredding the _W_ in somatic tissues, or by spreading some other allele that harms females only. If this female-harming allele were capable of gene drive in males, or were continually released into the wild, it could perhaps reach high enough frequencies to suppress the population. The _W_-shredder could be designed to also cause gene drive in males. Male gene drive could be accomplished using 'standard' CRISPR-Cas9 gene conversion, whereby the driving _Z_ allele would convert the wild type locus using homing endonuclease activity followed by homology-directed repair, causing heterozygous males to produce mostly drive-carrying sperm. Esvelt et al. note that male gene drive might not be necessary, since a _Z_-linked locus that prevents transmission of the _W_ may already enjoy a transmission advantage (Figures 1B-1C).

Here, I present an evolutionary simulation that can accommodate all of these hypothetical _Z_-linked drives. I aimed to test which properties of the gene drive and the ecology of the target species are critical to determining the likelihood and speed of extinction. For example, the gene drive will presumably spread faster if it can bias transmission in both sexes, but perhaps a simpler female-only drive would be perfectly adequate. Also, since the population will become more male-biased as the gene drive invades, there will be eco-evo feedback that might affect the evolutionary outcome in non-intuitive ways. For example, the altered sex ratio might intensify the fitness advantage accruing to any resistant _W_ chromosomes or autosomal modifiers that prevent _W_-shredding (due to Fisherian selection for an even sex ratio; [@holman2015co]), relative to that observed in earlier models focusing on gene drives carried on autosomes [@unckless2017ev; @drury2017cr]. Moreover, the change in sex ratio could affect the ecology and evolution of the population, particularly if males and females contribute differentially to density-dependent population growth [@rankin2007ma; @li2019int], or have different dispersal rates [@li2019sex]. The model incorporates the possibility that _Z_-linked resistant-to-drive alleles are sometimes created by NHEJ in heterozygote males, to test whether resistance is just as problematic as for autosomal drives [@gantz2015mu; @gantz2015hi; @hammond2016cr; @wang2016cr; @unckless2017ev].


# Methods

A full description of the simulation is provided as Supplementary Material. In brief, I simulate a finite population of dioecious diploids with _ZW_ sex determination, living in $k$ discrete habitat patches that are arranged linearly in a ring, and examine the demographic and evolutionary consequences of a single release of $n_{release}$ homozygote males carrying a _Z_-linked gene drive allele, _Z\*_. The drive allele causes biased inheritance and/or reduced fecundity in females, and optionally also causes biased inheritance in heterozygous males (e.g. via gene conversion). Each generation proceeds in discrete steps: birth, dispersal between patches, breeding within patches, and death of the parental generation. The equilibrium population size was roughly 10,000 in all simulations upon release of the gene drive, and the main outcomes of interest are the likelihood and speed of extinction. The simulation was written in R 3.4.0 and run on the Spartan cluster at the University of Melbourne; Table 1 lists the simulation parameters that were manipulated to study their effects.

Each male carries two autosomal loci (termed _A/a_ and _B/b_) and one _Z_-linked locus, while females carry both autosomal loci, a single allele at the _Z_-linked locus, plus a _W_ chromosome. There are three possible _Z_-linked alleles: the drive allele (_Z\*_), a wild-type allele ($Z^+$) that is vulnerable to gene drive in _Z\*_$Z^+$ males, and a resistant allele ($Z^r$) that is immune to gene drive in _Z\*_$Z^r$ males. Similarly, there are two types of _W_ chromosome: a wild-type _W_ that is vulnerable to shredding by the _Z\*_ allele ($W^+$), and an immune variant ($W^r$). The alleles _A_ and _B_ are dominant 'trans-acting' resistance alleles that confer immunity to _W_-shredding and gene conversion, respectively. The _Z\*_ allele imposes a cost $c_f$ on the fecundity of female carriers, and a cost $c_m$ on the mating success of male carriers. The resistance alleles $W^r$, $Z^r$, _A_ and _B_ are assumed to be cost-free. Setting $c_f = 1$ allows simulation of a female-sterilising _Z_-linked drive (Figure 1D).

Females carrying _Z\*_ (and no _A_ or $W^r$ alleles) produce $\frac{1}{2}(1 + p_{shred})$ _Z_-bearing gametes and $\frac{1}{2}(1 - p_{shred})$ _W_-bearing gametes, and thus produce mostly sons when $p_{shred} > 0$. Secondly, _Z\*_$Z^+$ males produce $\frac{1}{2}(1 + p_{conv} - p_{conv} p_{nhej})$ gametes carrying the _Z\*_ allele, $\frac{1}{2}(1 - p_{conv})$ gametes carrying the $Z^+$ allele, and $\frac{1}{2}(p_{conv} p_{nhej}))$ gametes carrying the $Z^r$ allele. Thus, gene conversion occurs in males if $p_{conv} > 0$, meaning that the _Z\*_ allele is over-represented in the gametes of these three male genotypes. The parameter $p_{nhej}$ represents the creation of resistance alleles via non-homologous end joining.

Female fecundity depends on the local and/or global density and fitness of other females, and the density of males, via functions involving five parameters that control the carrying capacity ($K$), maximum possible fecundity ($r$), the shape of the density-dependence function ($\alpha$), the relative contribution of males and females to density ($\delta$), and the scale of density-dependence (local or global; ($\psi$)). These parameters allow the model to capture a range of possible life histories suitable for different organisms that could be suppressed with _W_-shredders. For example, maximum fecundity $r$ is lower for birds than for cane toads, and $\delta$ might be lower in Lepidoptera than trematodes, since male trematodes consume space/resources inside the host (and males are much larger than females), while male Lepidoptera do not consume one of the main limitied resources that females need (i.e. host plants on which to lay their eggs). Female and male offspring disperse to other patches with probabilities $x_f$ and $x_m$ respectively, allowing for variable and sex-specific gene flow between patches. Dispersal was either local or global (i.e. to a neighbouring patch or a random patch).



# Results

## Three illustrative simulation runs
Figure 2 shows three contrasting simulation runs. In Figure 2A, the release of 20 _Z\*Z\*_ males at generation 50 resulted in invasion of the _Z\*_ allele, causing rapid extinction due to a lack of females. This simulation run assumed that the _Z\*_ alleles causes complete _W_-shredding, that _Z\*_ has minimal fitness costs, and there is no resistance to _W_-shredding (Table S3). 

In Figure 2B, _Z\*_ invaded but failed to cause extinction, even though it was assumed that $p_{shred} = 1$ and _W_-shredding was not resistable. However, this simulation did assume that individuals carrying at least one _Z\*_ allele paid heavy fitness costs ($c_f = 0.5$ and $c_m = 0.2$), and that there was no gene drive in males ($p_{conv} = 0$). The assumptions $p_{shred} = 1$ and $c_f = 0.5$ could imply that the _W_-bearing eggs/offspring of _Z\*_$W^+$ females are destroyed and not replaced, such that _W_-shredding increases the proportion but not the absolute number of offspring that inherit the _Z\*_ allele. Essentially _Z\*_ spreads via 'spite' [@gardner2006sp], in that it removes _W_ chromosomes from the local population and thereby makes room for more _Z\*_ alleles, creating indirect fitness benefits. However, the net fitness returns of the _Z\*_ allele's 'strategy' (i.e. sacrificing 20% fitness in males in order to remove _W_ chromosomes in females) decline as females become rarer, halting the spread of _Z\*_.

Lastly, Figure 2C shows a case where the invasion of _Z\*_ was reversed by the evolution of autosomal and _Z_-linked resistance alleles. Following the introduction of the _Z\*_ allele, resistant $Z^r$ mutants were created via non-homologous end joining, and then $Z^r$ spread to fixation due to its immunity to gene conversion in males. The autosomal resistance allele _A_ also invaded; _A_ confers resistance to _W_-shredding and was initially present in the population at 5% frequency. The spread of _A_ caused the sex ratio to revert to normal, preventing extinction, and _Z\*_ went extinct due to its direct fitness costs no longer being outweighed by the benefits of _W_-shredding and gene conversion. Incidentally, the resistant allele _A_ was favoured over _a_ because the male-biased population sex ratio created by _Z\*_ favours the production of daughters, and _AA_ and _Aa_ females produce more daughters than _aa_ females in populations where _Z\*_ is present.


## Effects of each parameter on the evolution of a _W_-shredder
Figure 3 shows the effects of each parameter for simulations of a _W_-shredder that potentially also benefits from gene drive in _Z\*Z_ males. Figure 4 shows the importance of each main effect and two-way interaction term to the extinction probability, while Figure S1 shows the effect of each parameter on the number of generations until extinction. Under favourable assumptions, extinction occurred around 20 generations after releasing _Z\*_, though it often took longer (Figure S1). Tables S1-S2 give the relative frequencies of the various possible outcomes (e.g. extinction of the population, or loss of _Z\*_).

In Figure 3, the parameters are arranged in order of their importance to extinction probability (see also Figure 4). By far the most important predictors of extinction were the efficiency of _W_-shredding in females ($p_{shred}$) and the existence of resistance against _W_-shredding: extinction never occurred unless $p_{shred}$ was high and autosomal alleles conferring resistance to _W_-shredding (allele _A_ in the model) were absent. This makes sense because a _W_-shredder cannot cause extinction unless _Z\*_-carrying females produce a strongly male-biased sex ratio and resistance to _W_-shredding cannot readily evolve. Extinction also occurred more quickly when $p_{shred}$ was 1 rather than 0.95 (Figure S1), further highlighting efficient _W_-shredding as an important design consideration.

The strength of gene drive in _Z\*Z_ males ($p_{conv}$; colours in Figure 3) also predicted extinction probability. However, $p_{conv}$ was less important than $p_{shred}$, and the _W_-shredder frequently caused extinction even when it showed normal Mendelian inheritance in males, or if resistance to male gene drive was common. The effect of male gene drive on extinction depended on other factors in the model (Figures 3, 4 and S2); for example, male gene drive was at its most beneficial when resistance to it could not evolve (either through pre-existing genetic variation, or the creation of resistant $Z^r$ alleles through NHEJ). Although its effects on extinction probability were somewhat small, male gene drive did hasten extinction considerably (Figure S1). For example, assuming perfect _W_-shredding, adding male gene drive with $p_{conv} = 0.95$ reduced the expected time to extinction from around 75 to 22 generations. 

The cost of the _Z\*_ allele to female fitness also affected extinction probability, and its effect interacted with the strength of gene drive in _Z\*Z_ males. Specifically, assuming that the _Z\*_ allele halves female fitness ($c_f = 0.5$) negates the fitness benefits of segregation distortion for the _Z\*_ allele, and so extinction could only occur when $c_f = 0.5$ if there was gene drive in males. Reassuringly, increasing $c_f$ from 0.01 or 0.1 had almost no effect on the likelihood of extinction, meaning that _W_ shredders might be an effective means of population control even if females carrying the gene drive suffer a 10% fitness cost. Similarly, assuming that _Z\*_ was costly to male carriers had little effect on extinction probability: extinction occurred almost as frequently when the reduction in male mating success was 20% rather than 1%. Both $c_f$ and $c_m$ were positively correlated with the time to extinction, particularly when there was no gene drive in males (Figure S1).

Some of the ecological variables also affected extinction probability. Chief among these was the shape parameter of the density-dependence function, $\alpha$. Setting $\alpha < 1$ causes female fecundity to decline at a decelerating rate with increasing population density, such that per-female fecundity only approaches its maximum value when the population is heavily depleted, making extinction more likely. Conversely for $\alpha > 1$, fecundity declines at an accelerating rate with increasing density, making extinction less likely due to the immediate increases in female fecundity that manifest once the population begins to shrink. Unsurprisingly, populations in which females have a higher maximum possible fecundity ($r$) were less likely to go extinct. Also, extinction was slightly more probable when female fecundity was determined by local density more than global density ($\psi$). This is because local density can remain high (and thus, per-female fecundity can remain low) even in meta-populations that are declining due to the spread of the _Z\*_ allele in some of their sub-populations.

Extinction probability also increased with $\delta$, the parameter that determines how male density affects female fecundity. When $\delta$ is high, female fecundity is constrained from increasing as the drive allele spreads by the ever-increasing proportion of males, contributing to extinction. Conversely, lower values of $\delta$ mean that male numbers are relatively unimportant in determining female fecundity, making extinction less likely because the shortage of females created by the gene drive alleviates competition on the remaining females. This result highlights that it is worth considering the ecology and population dynamics of target species when designing suppression drives that eliminate one sex.

Populations that are split into many semi-isolated patches were more difficult to drive extinct than those comparatively free of spatial structure, though the effect on extinction rate was small. The likely reason is that a highly-structured population creates refuges from the gene drive allele. The frequency and sex bias in dispersal was relatively unimportant to extinction probability, though there was a slight tendency for higher dispersal rates to stave off extinction, presumably because dispersal allows recolonisation of patches emptied by the gene drive. Similarly, it did not matter whether dispersal carried individuals to any patch, or only to neighbouring patches. Finally, there was no effect of the release strategy, suggesting that it may be unnecessary to release a _W_-shredding gene drive across the species' entire range provided that there is gene flow between patches. An additional implication of this result is that we cannot expect _Z_-linked gene drives to remain confined to their release sites, as previously found for autosomal drives [@noble2018cu].

## Effects of each parameter on a female-sterilising _Z_ drive
I also used the model to examine the evolution of a _Z_-linked allele that causes gene drive in males and also causes total sterility in females ($c_f = 1$; Figures S3-S6). This alternative type of gene drive was also effective at causing extinction, but only under the assumption that the population has little or no resistance to gene drive in males. For example, extinction never occurred if even 1% of the progeny of _Z\*Z_ males inherited a resistant $Z^r$ allele created by non-homologous end joining [c.f. @unckless2017ev]. Extinction also required that gene drive in males was strong (high $p_{conv}$), and that there were no autosomal resistance alleles to male gene drive. The effects of the other parameters in the model were similar as for a _W_-shredder, and extinction (when it occurred) took a fairly similar number of generations (around 25-30). 

## Interactions between model parameters
Many of the model parameters interacted in their effects on extinction probability (Figures 4 and S2). For _W_-shredders, increasing $p_{shred}$ only increased extinction probability provided that resistance to _W_-shredding was absent from the population, reaffirming the importance of resistance. Male gene drive was most beneficial when _Z\*W_ females had half the fecundity of wild types (i.e. $c_f = 0.5$) and when $p_{shred}$ was high, but male gene drive made little difference when $c_f \le 0.1$ or $p_{shred}$ was low. The demographic parameters $\alpha$ and $r$ were important to extinction rate only when $p_{shred} \le 1$; for $p_{shred} = 1$, the _W_-shredder was likely to cause extinction regardless of the ecological assumptions. For female-sterilising _Z_ drives, the most important interaction terms underscored the importance of efficient and unresistable male gene drive (Figures S5-S6).



# Discussion

The model shows that _W_-shredders are, in principle, very effective at eliminating populations, especially if _Z\*W_ females produce no daughters ($p_{shred} = 1$) and resistance to _W_-shredding cannot evolve. The results have implications for the design of _Z_-linked _W_-shredders and female-sterilising suppression drives.

One design consideration is whether to engineer _W_-shredders that are also capable of gene drive in males, e.g. by including guide RNAs that target the _Z_ as well as the _W_ chromosome. In the model, _W_-shredders very often caused extinction even without male gene drive (i.e. when $p_{conv} = 0$), provided that females carrying the _W_-shredder had comparable fecundity to wild type females, and that carrier females produce very few daughters (as in Figure 1B). Conversely if _W_-shredder females had low fecundity (around half that of a wild type, or below; Figure 1C) or produced some daughters, male gene drive was often essential for the _W_-shredder to cause extinction, or at least for extinction to occur rapidly enough to be useful. Although male gene drive was not always essential to extinction, it did reduce the number of generations until extinction occurred, sometimes substantially. Therefore, I conclude that it would almost certainly be worth the effort to incorporate a male-acting gene drive if developing a _W_-shredder for species with long generation times, such as invasive birds. Conversely, the rate of population decline may be adequate even without male gene drive for species that have multiple generations per year, such as Lepidoptera and _Schistosoma_ parasites. Foregoing male drive could simplify the design of _W_-shredders since they would only need to target the _W_ chromosome (and not also the _Z_), particularly because male-acting gene conversion drives seem more challenging to develop than female-acting ones in some taxa (due to sex differences in DNA repair; [@grunwald2019super]). Conversely, strong male gene drive was always essential to extinction for female-sterilising suppression drives (Figure 1D). _Z_-linked alleles that drive in males and cause sterility in females were effective at causing extinction, but were very vulnerable to the evolution of resistance to male gene drive (e.g. via drive-resistant alleles created by NHEJ; [@unckless2017ev]). 

Another aim when designing _W_-shredders should be to ensure that female carriers produce as few daughters as possible (ideally none), while producing a large number of drive-carrying sons (ideally as many as the total offspring produced by non-carriers). This implies that one should ideally design a construct that cleaves the _W_ chromosome early in gametogenesis or development, to increase the chance that the number of surviving progeny produced by each female is unaffected. Cleavage of the _W_ should also be restricted to the female germ line, to minimise fitness losses due to the loss of the _W_ in somatic cells. For some species, this may be as simple as placing the _W_-shredder under the control of a promoter such as _nanos_ [@champer2018re; @zhang2018si], assuming that females are able to replace lost _W_-bearing oocytes before they are provisioned with limiting resources. Even if the lost daughters are not replaced with sons, the _Z\*_ allele might still exhibit drive because the surviving _Z\*_ sons will experience reduced competition (somewhat like _Medea_ [@hay2010en]). In Lepidoptera, juvenile density is often strongly negatively correlated with survival, and there are various maternally-transmitted endosymbionts that drive through populations by killing males to lessen competition on their infected sisters (e.g. [@jiggins2000bu; @jiggins2003ma]); these observations suggest that _W_-shredder alleles might invade Lepidopteran populations even if _Z\*W_ females produced half as many viable eggs, though male gene drive would certainly help the invasion.

The _W_-shredding mechanism should also be designed in a way that makes it difficult for _W_-linked or _trans_-acting resistance to shredding to evolve. One way to do this would be to use a single guide RNA that targets high copy number _W_-specific sequences, or to use multiple guide RNAs that target multiple _W_-linked sequences [@champer2018re; @oberhofer2018be]. This way, multiple changes to the reference sequence would be required for a _W_ chromosome to acquire resistance to cleavage by the _W_-shredder. To ensure that the targets of cleavage do not become resistant as a result of indels induced by NHEJ, one can ensure that the guide RNA's target lies within an essential gene where an indel would be selectively disadvantageous, preventing resistant alleles from accumulating in the population [@champer2018re; @oberhofer2018be]. This may not be necessary if the _W_-shredder targets many _W_-linked loci, but it is an important design consideration for the male component of the gene drive, because the evolution of _Z_-linked resistance completely nullified the usefulness of male gene drive in the simulation (echoing [@unckless2017ev]). Recent work demonstrated the feasibility of arrays containing many guide RNAs separated by spacers [@kurata2018hi], suggesting it may soon be easier to create gene drives with multiple guide RNAs.

The model also indicated that extinction does not require the release of large numbers of individuals: releasing just 20 _Z\*_ males was often enough to eliminate a spatially-structured metapopulation of 10,000 individuals in a few generations. On the one hand, this is advantageous because _W_-shredders would be cheap and easy to deploy once they are developed, and they are likely to extirpate whole metapopulations even if gene flow is weak. However, such high invasiveness is not always desirable, because it makes the gene drive more difficult to restrict to a particular area. This could limit the usefulness of _W_-shredders to control species like Lepidoptera and birds, where one may wish to eradicate only invasive or agriculturally damaging populations, while leaving other populations untouched. Modifications to gene drive design -- such as the self-limiting 'daisy drive' system -- are being developed to address this important concern [@min2017da; @noble2019da].

The model further showed that _W_-shredders can fail to cause extinction if carrier individuals have low fitness, although extinction was frequently observed even if these fitness costs were substantial. Populations in which females can become highly fecund as the population shrinks (i.e. low $\alpha$ and high $r$) were also less likely to go extinct, though extinction tended to occur anyway provided $p_{shred} = 1$. The model also highlighted that _W_-shredders, and indeed any gene drive that creates a male-biased sex ratio, are most effective in suppressing species in which the density of males is an important determinant of population growth, e.g. because males use resources that females need [@li2019int]. By contrast, if male density is not very important to population growth (e.g. because females are limited by a resource that is not consumed by males), female fecundity increases as females become rarer, slowing the decline in population size caused by the _W_-shredder and potentially staving off extinction. Interestingly, the sexes are very different in the _Schistosoma_ trematodes responsible for schistosomiasis, which have been proposed as candidates for control using a _W_-shredder by Kevin Esvelt and colleagues. Female _Schistosoma_ live inside the body of the much larger male, who feeds on the host's blood and passes some of it to the female. Presumably, this means that the number of males (not females) is the primary determinant of whether a host or habitat is saturated, making _Schistosoma_ a good candidate for control with _W_-shredders. In Lepidoptera and birds -- two other _ZW_ taxa that could potentially be controlled with _W_-shredders -- males and females generally have similar ecological niches, such that _W_-shredders should be effective. Other ecological parameters like the patchiness of the population ($k$), the frequency and sex bias of dispersal ($x_f$ and $x_m$), and the scale of competition ($\psi$) had relatively little effect on the probability of extinction.

Finally, I note that _W_-shredders might in general be easier to develop than _X_-shredders. Efforts to develop an _X_-shredder in _Anopheles_ mosquitos were initially hindered because the I-PpoI protein used to cleave the _X_ was paternally transmitted to the embryo inside sperm, causing all embryos to die (not just daughters) due to loss of the maternally-inherited _X_. Although this technical issue was later mitigated [@galizi2014sy], such intergenerational effects would not trouble a _W_-shredder since the _W_ chromosome is unique to females (provided that the _W_-shredding protein was not expressed in males and/or was not transmitted in their sperm). Additionally, _W_-shredders might sometimes be easier to develop than gene drives that work by deleting genes that are essential to female (but not male) fitness [e.g. @burt2018se]. This is because one could design a prototype _W_-shredder based only on sequence data from the sex chromosomes, while identifying genes with female-specific fitness effects requires more detailed data (e.g. expression profiling or knockout studies) that are unavailable for some taxa.


\newpage
# Tables

**Table 1**: List of variables, and their corresponding parameter(s) in the model, which were varied in order to study their effects on extinction.
```{r xtable, results="asis"}
print_table1()
```


\newpage
# Figures

\begin{figure}[h]
\centering
\includegraphics[width=1.0\textwidth]{../figures/figure1_moths.pdf}
\caption{\footnotesize{Some hypothetical \textit{Z}-linked suppression drives considered in this study. Panel A illustrates normal inheritance of sex chromosomes in a wild type \textit{ZW} female (assumed to be mated to a wild type \textit{ZZ} male; not shown): the offspring sex ratio is even. In panel B, the female carries a \textit{W}-shredder allele ($Z^*$) that kills gametes or offspring early enough that missing daughters are replaced with more $Z^*$-bearing sons. In panel C, the lost daughters are not replaced, though their absence increases the survival probability of the sons somewhat (shown by their larger size), causing super-Mendelian inheritance of the $Z^*$ allele. Lastly, panel D shows a \textit{Z}-linked female-sterilising allele (e.g. an allele that cleaves the \textit{W} chromosome or a female-essential gene in somatic cells); since it is strongly disadvantageous in females, such an allele would go extinct unless it benefits from gene drive in heterozygous males.}}
\end{figure}
\newpage

\begin{figure}[h]
\centering
\includegraphics[width=1.0\textwidth]{../figures/figure2.pdf}
\caption{\footnotesize{Three illustrative runs of the simulation, showing evolution in response to the introduction of 20 males carrying a \textit{W}-shredder at Generation 50 (dotted line). In panel A, the driving $Z^*$ allele fixed very quickly, causing population extinction through a shortage of females. In panel B, the $Z^*$ allele spread until its fitness costs began to negate its transmission advantage, causing the population to persist at a reduced size. In panel C, the $Z^*$ allele invaded, which selected for the resistance alleles \textit{A} and $Z^r$ and caused $Z^*$ to go extinct. The population size \textit{N} is shown as a fraction of its maximum value of 10,000. Table S3 gives the parameter spaces used for these three runs.}}
\end{figure}
\newpage

\begin{figure}[h]
\centering
\includegraphics[width=1.0\textwidth]{../figures/figure3_inkscape.eps}
\caption{\footnotesize{The percentage of simulations of a \textit{W}-shredder that ended in extinction, for all runs with a particular value (shown on the \textit{x}-axis) for a given parameter (shown in the panels). For example, there were no extinctions in any of the thousands of runs for which I assumed $p_{shred} = 0.5$, while 60\% of runs where $p_{shred} = 1$ resulted in extinction. The panels are ordered by the range of the \textit{x}-axis, which indicates the relative importance of each variable to extinction probability. Figure S3 gives a similar plot for simulations of a female-sterilising $Z^*$ allele.}}
\end{figure}
\newpage

\begin{figure}[h]
\centering
\includegraphics[width=1.0\textwidth]{../figures/figure4_inkscape.pdf}
\caption{\footnotesize{Relative parameter importance in the simulations of \textit{W}-shredders, for the top 25 most important main effects or two-way interactions (from a binomial GLM that included all the main effects and all their two-way interactions). Each predictor variable was scaled before running the model, meaning that the absolute effect size indicates how important each parameter is to the extinction probability, given the range of values plotted in Figure 3. Figure S5 gives a similar plot for simulations of a female-sterilising $Z^*$ allele.}}
\end{figure}
