---
title: "plot_model"
author: "Luke Holman"
date: "2018-07-17"
output: workflowr::wflow_html
editor_options:
  chunk_output_type: console
---

```{r setup, echo=FALSE, results='hide'}
source_rmd <- function(file, local = FALSE){
  options(knitr.duplicate.label = "allow")
  tempR <- tempfile(tmpdir = ".", fileext = ".R")
  on.exit(unlink(tempR))
  print(file.exists("model_functions.Rmd"))
  knitr::purl(file, output = tempR, quiet = TRUE)
  source(tempR, local = globalenv())
}
source_rmd("analysis/model_functions.Rmd")
```

# Load packages and results

```{r}
library(ggplot2)
# library(ggbeeswarm)
library(ggthemes)
library(gridExtra)
library(biglm)
# library(brms)
library(ggeffects)
library(readr)
library(gridExtra)
library(grid)
library(pander)
library(latex2exp)
library(Cairo)
library(car)
library(dplyr)
library(purrr)
library(reshape2)


if(!file.exists("data/all_results.rds")){
  results <- lapply(list.files(path = "data", pattern = "results_", full.names = T), readRDS) %>% 
    do.call("rbind", .) %>%
    filter(cost_A == 0 & cost_B == 0)
  saveRDS(results, file = "data/all_results.rds")
} else results <- read_rds("data/all_results.rds")

results$went_extinct <- ifelse(results$outcome == "Population extinct", 1, 0)

results$migration_type[results$migration_type == "local"] <- "Local"
results$migration_type[results$migration_type == "global"] <- "Global"

find_variable_parameters <- function(dat){
  dat %>% 
    select(-id, -realisations, -generation_extinct, -generation_Zd_extinct, 
           -generation_W_extinct, -generation_Zd_fixed, -outcome, -went_extinct, 
           -mating_table, -initial_Wr, -initial_Zr) %>%
    sapply(function(x) length(unique(x))) %>% 
    keep(~.x > 1) %>% names()
}

variable_parameters <- find_variable_parameters(results)
combinations <- apply(combn(variable_parameters, 2), 2, paste0, collapse = " x ")

nice_names <- data.frame(original = c(variable_parameters, combinations),
                         new = gsub("_", " ", c(variable_parameters, combinations)),
                         stringsAsFactors = FALSE) %>%
  mutate(
    new = gsub("rel", "Rel", new),
    new = gsub("W shredding rate", "Gene drive in females ($p_{shred}$)", new),
    new = gsub("Z conversion rate", "Gene drive in males ($p_{conv}$)", new),
    new = gsub("Zr creation rate", "Rate of NHEJ ($p_{nhej}$)", new),
    new = gsub("Zr mutation rate", "Z-linked resistance ($\\mu{_Z}$)", new),
    new = gsub("Wr mutation rate", "W-linked resistance ($\\mu{_W}$)", new),
    new = gsub("cost Zdrive female", "Cost of Z* to females ($c_f$)", new),
    new = gsub("cost Zdrive male", "Cost of Z* to males ($c_m$)", new),
    new = gsub("male migration prob", "Male dispersal frequency", new),
    new = gsub("feMale dispersal frequency", "Female dispersal frequency (xf)", new),
    new = gsub("Male dispersal frequency", "Male dispersal frequency (xm)", new),
    new = gsub("xm", "$x_m$", new), 
    new = gsub("xf", "$x_f$", new),
    new = gsub("migration type", "Dispersal type", new),
    new = gsub("n patches", "Number of patches ($k$)", new),
    new = gsub("softness", "Importance of local density ($\\psi$)", new),
    new = gsub("male weighting", "Male effect on density ($\\delta$)", new),
    new = gsub("density dependence shape", "Shape of density dependence ($\\alpha$)", new),
    new = gsub("max fecundity", "Maximum fecundity ($r$)", new),
    new = gsub("initial A", "W-shredding resistance allele A present", new),
    new = gsub("initial B", "Gene conversion resistance allele B present", new),
    new = as.character(TeX(new))) %>% 
  mutate(new = gsub("mu", "\\mu", new))

# Sort the results between those that model a W-shredder, versus those that model a Z* that sterilisies females
W_shredder <- results %>% filter(cost_Zdrive_female != 1)
female_sterilising <- results %>% filter(cost_Zdrive_female == 1)
rm(results)
variable_parameters_W_shredder <- find_variable_parameters(W_shredder)
variable_parameters_female_sterilising <- find_variable_parameters(female_sterilising)
```

## Table showing the frequencies of each possible outcome
```{r}
make_tally_table <- function(dat){
  outcomes <- dat$outcome %>% table() %>% melt() %>% arrange(-value) %>% mutate(p = round(100 * value / sum(value), 1))
  names(outcomes) <- c("Outcome", "Number of simulations", "%")
  outcomes %>% mutate(Outcome = gsub("without", "without causing", Outcome),
                      Outcome = gsub("n extinct", "n went extinct", Outcome),
                      Outcome = gsub("d extinct", "d went extinct", Outcome)) %>% pander()
}

make_tally_table(W_shredder)
```


```{r}
make_tally_table(female_sterilising)
```

## Make Figure 1

```{r}
a <- read_rds("data/allele_freqs_1.rds")

plot_run <- function(){
  
  get_data <- function(model_id, allele_freqs_list, result_df, label){
    
    df <- pluck_allele_freqs(model_id, allele_freqs_list) 
    
    alleles_to_plot <- group_by(df, allele) %>% 
      summarise(uniques = length(unique(frequency))) %>% 
      filter(uniques > 1) %>% # Don't plot alleles that stay at 0 whole time
      pull(allele)
    
    alleles_to_plot <- c(alleles_to_plot, "N")
    df <- df[df$allele %in% alleles_to_plot, ]
    df$frequency[df$allele == "N"] <- df$frequency[df$allele == "N"] /
      max(df$frequency[df$allele == "N"])
     
    
    paras <- result_df %>% mutate(id = as.character(id)) %>% filter(id == model_id)
    last_generation <- tail(df,12) %>% select(allele, frequency)
    print(paras); print(last_generation)
    
    df %>% mutate(facet = label,
                  allele = replace(allele, allele == "Zd", "Z*"),
                  allele = replace(allele, allele == "females", "Females"))
  }
  
  rbind(
    get_data("10024119427447", a, W_shredder, "A. Population extinction"),
    get_data("1002596940286", a, W_shredder, "B. Failure to cause extinction"),
    get_data("10024112958792", a, W_shredder, "C. Resistance prevents extinction")
  ) %>%
    ggplot(aes(x = generation, y = frequency, colour = allele, group = allele)) + 
    geom_vline(xintercept = 50, linetype = 2, colour = "grey10", size = 0.9) + 
    geom_line(size = 0.9, alpha = 0.9) + 
    facet_wrap(~facet, scales = "free_x") +
    theme_hc() + scale_colour_hc(name = "") +
    theme(strip.text = element_text(hjust = 0), 
          strip.background = element_rect(fill = "grey90"),
          legend.position = "top") + 
    xlab("Generation") + ylab("Frequency")
}

fig1 <- plot_run()
cairo_pdf(file = "figures/figure1.pdf", width = 10, height = 10); fig1; dev.off()
fig1
```
<br></br>
**Figure 1:**



## Make Figure 2

```{r}
get_percent_extinct <- function(dat, parameter){
  dat %>%
    group_by(!! sym(parameter)) %>%
    summarise(extinct = sum(went_extinct),
              not_extinct = n() - sum(went_extinct),
              prob = list(binom.test(extinct, extinct + not_extinct))) %>%
    rowwise() %>%
    mutate(percent_extinct = 100 * extinct/(extinct + not_extinct),
           lower_95_CI = 100 * prob$conf.int[1],
           upper_95_CI = 100 * prob$conf.int[2],
           parameter =  parameter) %>% rename(value = !! sym(parameter)) %>%
    select(parameter, value, everything()) %>% select(-prob) %>% ungroup()
} 

W_shredder_percent_extinct <- lapply(variable_parameters, get_percent_extinct, dat = W_shredder) %>% do.call("rbind", .)
female_sterilising_percent_extinct <- lapply(variable_parameters, get_percent_extinct, dat = female_sterilising) %>% do.call("rbind", .)

make_figure_2 <- function(percent_extinct_data){
  parameter_importance <- percent_extinct_data %>%
    mutate(parameter = nice_names$new[match(parameter, nice_names$original)]) %>%
    group_by(parameter) %>% summarise(range = max(percent_extinct) - min(percent_extinct)) %>%
    arrange(-range) %>% pull(parameter)
  
  levels <- levels(factor(percent_extinct_data$value))
  levels <- c(levels[!(levels %in% c("all_patches", 100))], 100, "all_patches")
  levels <- replace(levels, levels == "all_patches", "Release in\nall patches")
  levels <- replace(levels, levels == "one_patch", "Release in\na single patch")
  
  percent_extinct_data %>% 
    mutate(parameter = nice_names$new[match(parameter, nice_names$original)],
           parameter = factor(parameter, parameter_importance),
           value = replace(value, value == "all_patches", "Release in\nall patches"),
           value = replace(value, value == "one_patch", "Release in\na single patch"),
           value = factor(value, levels)) %>%
    ggplot(aes(value, percent_extinct)) + 
    geom_errorbar(aes(ymin=lower_95_CI, ymax = upper_95_CI), width = 0) +
    geom_point() + 
    facet_wrap(~parameter, scales = "free", labeller = label_parsed) + 
    xlab("Parameter value") +
    ylab("% simulation runs resulting in extinction") + 
    theme_bw() + 
    theme(strip.background = element_blank(),
          panel.grid.major.x = element_blank())
  
}

fig2 <- make_figure_2(W_shredder_percent_extinct) 
cairo_pdf(file = "figures/figure2.pdf", width = 10, height = 10); fig2; dev.off()
fig2
```
<br></br>
**Figure 2:**


## Make Figure SX

```{r}
make_figure_2(female_sterilising_percent_extinct)
```
<br></br>
**Figure SX:**


## Finding variables with interacting effects on the extinction probability


### Testing for interactions using GLM 
Run a binomial generalized linear model (GLM), with all of the variable model parameters and all their 2-way interactions as predictors. The response variable is extinction, coded as a zero or one. Since the number of data points is very large (millions), I use the `biglm` package ("big GLM"). I run a model separately on all the runs that considered the evolution of a _W_-shredder (i.e. females are viable, but produce mostly sons), and the run that considered a female-sterilising _Z\*_.  

```{r}
run_big_glm <- function(dat, sterilising = FALSE){
  
  if(sterilising){
    variable_parameters <- variable_parameters[!(variable_parameters %in% c("W_shredding_rate", "cost_Zdrive_female"))] 
  }
  
  formula <- paste("went_extinct ~ (",
                   paste0(variable_parameters, collapse = " + "), ")^2",
                   sep = "")
  
  if(!sterilising) formula <- paste(formula, "- W_shredding_rate:cost_Zdrive_female")
  
  my_scale <- function(x) as.numeric(scale(x))
  
  glm_model <- bigglm(as.formula(formula), data = dat %>% mutate_if(is.numeric, my_scale), chunksize = 10000)
  
  output <- data.frame(Parameter = rownames(summary(glm_model)[[2]]), summary(glm_model)[[2]]) %>% 
    as_tibble() %>% arrange(-abs(Coef)) %>%
    mutate(Parameter = gsub(":", " x ", Parameter),
           Parameter = factor(Parameter, rev(Parameter)),
           log10_p = -log10(p + 1e-300)) 
  
  names(output)[2:4] <- c("Estimate", "Lower_95_CI", "Upper_95_CI")
  output
}

if(!file.exists("data/W_shredder_model.rds")) {
  W_shredder_model <- run_big_glm(W_shredder)
  saveRDS(W_shredder_model, "data/W_shredder_model.rds")
  female_sterilising_model <- run_big_glm(female_sterilising, sterilising = TRUE)
  saveRDS(female_sterilising_model, "data/female_sterilising_model.rds")
} else {
  W_shredder_model <- readRDS("data/W_shredder_model.rds")
  female_sterilising_model <- readRDS("data/female_sterilising_model.rds")
}
```


### Finding the top-ranked effects in the GLM

```{r}
importance_plot <- function(dat, n = "all"){
  if(n != "all") dat <- dat %>%
      mutate(row=1:n()) %>% filter(row %in% 1:n)
  
  for(i in 1:nrow(dat)){
    if(dat$Estimate[i] < 0){
      lower <- dat$Lower_95_CI[i]
      dat$Lower_95_CI[i] <- dat$Upper_95_CI[i]
      dat$Upper_95_CI[i] <- lower
    }
  }
  
  dat <- dat %>%
    mutate(Parameter = gsub("one_patch", "", Parameter),
           Parameter = gsub("Local", "", Parameter),
           Parameter = nice_names$new[match(Parameter, nice_names$original)]) 
  
  dat %>%
    ggplot(aes(reorder(Parameter, abs(Estimate)), abs(Estimate))) + 
    geom_bar(stat = "identity", fill = "tomato") +
    geom_errorbar(aes(ymin = abs(Lower_95_CI), ymax = abs(Upper_95_CI)), width = 0) + 
    coord_flip() + 
    scale_y_continuous(expand = c(0, 0)) + 
    scale_x_discrete(labels = parse(text = dat$Parameter)) + xlab("Model parameter") + ylab("Absolute effect size")

}
importance_plot(W_shredder_model, n = 25)
```
<br></br>
**Figure SX:**

```{r}
importance_plot(female_sterilising_model, n = 25)
```
<br></br>
**Figure SX:**

### Plot showing interacting effects on extinction probability

```{r}
get_percent_extinct_double <- function(dat, p1, p2){
  output <-  dat %>%
    group_by(!! sym(p1), !! sym(p2)) %>%
    summarise(extinct = sum(went_extinct),
              not_extinct = n() - sum(went_extinct),
              prob = list(binom.test(extinct, extinct + not_extinct))) %>%
    rowwise() %>%
    mutate(percent_extinct = 100 * extinct / (extinct + not_extinct),
           lower_95_CI = 100 * prob$conf.int[1],
           upper_95_CI = 100 * prob$conf.int[2],
           parameter_1 =  p1, 
           parameter_2 =  p2) 
  names(output)[1:2] <- paste("value", 1:2, sep = "_")
  output %>%
    mutate(value_1 = as.character(value_1), value_2 = as.character(value_2)) %>%
    select(parameter_1, parameter_2, value_1, value_2, everything()) %>% 
    select(-prob) %>% ungroup()
} 

# combinations <- combn(variable_parameters, 2)
# two_way_W_shredder <- map2_df(combinations[1, ], 
#                               combinations[2, ], 
#                               get_percent_extinct_double, dat = W_shredder)

interesting <- c("W_shredding_rate~initial_A", "Z_conversion_rate~cost_Zdrive_female", "W_shredding_rate~Wr_mutation_rate",
                 "W_shredding_rate~Z_conversion_rate", "Zr_creation_rate~cost_Zdrive_female", "W_shredding_rate~density_dependence_shape",
                 "cost_Zdrive_female~n_patches", "W_shredding_rate~max_fecundity", "density_dependence_shape~max_fecundity")
two_way_W_shredder <- map2_df(str_split(interesting, "~", simplify = TRUE)[, 1], 
                              str_split(interesting, "~", simplify = TRUE)[, 2], 
                              get_percent_extinct_double, dat = W_shredder)


two_way_plot <- function(two_way_data, pairwise_combos){
  
  two_way_data <- two_way_data %>%
    mutate(pasted = paste(parameter_1, parameter_2, sep="~")) %>%
    filter(pasted %in% pairwise_combos) 

  z_range <- range(two_way_data$percent_extinct) * 1.1
  
  make_one <- function(df){
    xlab <- df$parameter_1[1]
    ylab <- df$parameter_2[1]
    ggplot(df, aes(value_1, value_2, fill = percent_extinct)) + 
      geom_tile(colour = "black") + 
      scale_fill_distiller(palette = "PuBuGn", direction = 1, limits = z_range, name = "% extinct") + 
      xlab(xlab) + ylab(ylab) + 
      theme_minimal() + 
      theme(panel.border = element_rect(fill = NA, colour = "black")) + 
      scale_x_discrete(expand = c(0,0)) + 
      scale_y_discrete(expand = c(0,0)) 
  }
  
  grid_arrange_shared_legend <- function(...) {
    plots <- list(...)
    g <- ggplotGrob(plots[[1]] + theme(legend.position="bottom"))$grobs
    legend <- g[[which(sapply(g, function(x) x$name) == "guide-box")]]
    lheight <- sum(legend$height)
    grid.arrange(
      do.call(arrangeGrob, lapply(plots, function(x)
        x + theme(legend.position = "none"))),
      legend,
      ncol = 1,
      heights = unit.c(unit(1, "npc") - lheight, lheight))
  }

  dat_list <- vector(mode = "list", length = length(pairwise_combos))
  for(i in 1:length(pairwise_combos)){
    dat_list[[i]] <- two_way_data %>% filter(pasted == pairwise_combos[i])
  }
  
  lapply(dat_list, make_one) %>% do.call("grid_arrange_shared_legend", .)
}

two_way_plot(two_way_W_shredder, interesting)
```
<br></br>
**Figure 3:**
