---
title: "Plotting the simulation results"
output: workflowr::wflow_html
editor_options:
  chunk_output_type: console
---

```{r setup, echo=FALSE, results='hide', message=FALSE, warning=FALSE}
source_rmd <- function(file, local = FALSE){
  options(knitr.duplicate.label = "allow")
  tempR <- tempfile(tmpdir = ".", fileext = ".R")
  on.exit(unlink(tempR))
  print(file.exists("model_functions.Rmd"))
  knitr::purl(file, output = tempR, quiet = TRUE)
  source(tempR, local = globalenv())
}
source_rmd("analysis/model_functions.Rmd") 
```

## Load packages and results

```{r results='hide', message=FALSE, warning=FALSE}
packages <- c("dplyr", "purrr", "ggplot2", "reshape2", "Cairo", "knitr",  
              "latex2exp", "pander", "grid", "gridExtra", "ggthemes", "data.table",
              "readr", "tibble", "biglm", "kableExtra", "future", "future.apply")
shh <- suppressMessages(lapply(packages, library, character.only = TRUE, quietly = TRUE))

if(!file.exists("data/all_results.rds")){
  
  # Open all the results files and bind them together into a tibble
  results <- lapply(list.files(path = file.path(getwd(), "data"), pattern = "results_", full.names = TRUE), 
                    function(focal_file) {
                      readRDS(focal_file) %>% filter(cost_A == 0 & cost_B == 0)}) %>% 
    rbindlist() %>% as_tibble()
  
  # Make sure each unique parameter space is run exactly once
  results <- results %>% 
    distinct(release_strategy, W_shredding_rate, Z_conversion_rate,
             Zr_creation_rate, Zr_mutation_rate, Wr_mutation_rate,
             cost_Zdrive_female, cost_Zdrive_male, male_migration_prob,
             female_migration_prob, migration_type, n_patches, softness,
             male_weighting, density_dependence_shape, max_fecundity, realisations,
             initial_A, initial_B, .keep_all = TRUE)
  
  saveRDS(results, file = "data/all_results.rds")
} else results <- read_rds("data/all_results.rds")

results$went_extinct <- ifelse(results$outcome == "Population extinct", 1, 0)
results$migration_type[results$migration_type == "local"] <- "Local"
results$migration_type[results$migration_type == "global"] <- "Global"

# Command to upload the results file to Spartan - needed for Spartan to queue up all the parameter spaces not already finished
# scp /Users/lholman/Rprojects/W_shredder/data/all_results.rds lukeholman@spartan:/data/projects/punim0243/W_shredder/data/all_results.rds

# Helper function to get a list of all the model parameters that vary between runs
find_variable_parameters <- function(dat){
  dat %>% 
    select(-id, -realisations, -generation_extinct, -generation_Zd_extinct, 
           -generation_W_extinct, -generation_Zd_fixed, -outcome, -went_extinct, 
           -mating_table, -initial_Wr, -initial_Zr) %>%
    sapply(function(x) length(unique(x))) %>% 
    keep(~.x > 1) %>% names()
}
variable_parameters <- find_variable_parameters(results)
combinations <- apply(combn(variable_parameters, 2), 2, paste0, collapse = " x ") # all 2-way combos of parameters

# Make a data frame to convert R-friendly names to figure-friendly names
nice_names <- data.frame(
  original = c(variable_parameters, combinations),
  new = gsub("_", " ", c(variable_parameters, combinations)),
  stringsAsFactors = FALSE) %>%
  mutate(
    new = gsub("rel", "Rel", new),
    new = gsub("W shredding rate", "Gene drive in females ($p_{shred}$)", new),
    new = gsub("Z conversion rate", "Gene drive in males ($p_{conv}$)", new),
    new = gsub("Zr creation rate", "Rate of NHEJ ($p_{nhej}$)", new),
    new = gsub("Zr mutation rate", "Z-linked resistance ($\\mu{_Z}$)", new),
    new = gsub("Wr mutation rate", "W-linked resistance ($\\mu{_W}$)", new),
    new = gsub("cost Zdrive female", "Cost of Z* to females ($c_f$)", new),
    new = gsub("cost Zdrive male", "Cost of Z* to males ($c_m$)", new),
    new = gsub("male migration prob", "Male dispersal frequency", new),
    new = gsub("feMale dispersal frequency", "Female dispersal frequency (xf)", new),
    new = gsub("Male dispersal frequency", "Male dispersal frequency (xm)", new),
    new = gsub("xm", "$x_m$", new), 
    new = gsub("xf", "$x_f$", new),
    new = gsub("migration type", "Dispersal type", new),
    new = gsub("n patches", "Number of patches ($k$)", new),
    new = gsub("softness", "Importance of local density ($\\psi$)", new),
    new = gsub("male weighting", "Male effect on density ($\\delta$)", new),
    new = gsub("density dependence shape", "Shape of density dependence ($\\alpha$)", new),
    new = gsub("max fecundity", "Maximum fecundity ($r$)", new),
    new = gsub("initial A", "Initial freq. W-shredding resistance allele A", new),
    new = gsub("initial B", "Initial freq. gene conversion resistance allele B", new),
    new = as.character(TeX(new))) %>% 
  mutate(new = gsub("mu", "\\mu", new))

# Sort the results between those that model a W-shredder, versus those that model a Z* that sterilises females
W_shredder <- results %>% filter(cost_Zdrive_female != 1)
female_sterilising <- results %>% filter(cost_Zdrive_female == 1)
rm(results)
```

## Table showing the frequencies of each possible outcome

**Table S1**: The number and percentage of simulation runs that ended with the five possible outcomes, for the subset of simulation runs focusing on a _W_-shredder gene drive.
<br></br>
```{r}
make_tally_table <- function(dat){
  outcomes <- dat$outcome %>% table() %>% melt() %>% arrange(-value) %>% mutate(p = round(100 * value / sum(value), 1))
  names(outcomes) <- c("Outcome", "Number of simulations", "%")
  outcomes %>% mutate(Outcome = gsub("without", "without causing", Outcome),
                      Outcome = gsub("n extinct", "n went extinct", Outcome),
                      Outcome = gsub("d extinct", "d went extinct", Outcome),
                      Outcome = gsub("Zd", "Z*", Outcome)) 
}

table_S1 <- make_tally_table(W_shredder) 
saveRDS(table_S1, file = "figures/tableS1.rds")
table_S1 %>% pander()
```

**Table S2**: The number and percentage of simulation runs that ended with the five possible outcomes, for the subset of simulation runs focusing on a female-sterilising _Z_-linked gene drive.
<br></br>
```{r}
table_S2 <- make_tally_table(female_sterilising)
saveRDS(table_S2, file = "figures/tableS2.rds")
table_S2 %>% pander()
```

## Make Figure 2

```{r results='hide'}
plot_run <- function(){
  
  get_data <- function(model_id, allele_freqs_list, result_df, label){
    
    df <- pluck_allele_freqs(model_id, allele_freqs_list) 
    
    alleles_to_plot <- group_by(df, allele) %>% 
      summarise(uniques = length(unique(frequency))) %>% 
      filter(uniques > 1) %>% # Don't plot alleles that stay at 0 whole time
      pull(allele)
    
    alleles_to_plot <- c(alleles_to_plot, "N")
    df <- df[df$allele %in% alleles_to_plot, ]
    df$frequency[df$allele == "N"] <- df$frequency[df$allele == "N"] /
      max(df$frequency[df$allele == "N"])
     
    paras <- result_df %>% mutate(id = as.character(id)) %>% filter(id == model_id)
    last_generation <- tail(df,12) %>% select(allele, frequency)
    print(paras); print(last_generation)

    df %>% mutate(facet = label,
                  allele = replace(allele, allele == "Zd", "Z*"),
                  allele = replace(allele, allele == "females", "Females"))
  }
  
  allele_freqs <- read_rds("data/allele_freqs_1.rds")
  
  rbind(
    get_data("10024119427447", allele_freqs, W_shredder, "A. Z* causes rapid extinction"),
    get_data("1002596940286", allele_freqs, W_shredder, "B. Z* fails to cause extinction"),
    get_data("10024112958792", allele_freqs, W_shredder, "C. Resistance to Z* prevents extinction")) %>%
    ggplot(aes(x = generation, y = frequency, colour = allele, group = allele)) + 
    geom_vline(xintercept = 50, linetype = 3, colour = "grey10", size = 0.9) + 
    geom_line(size = 0.9, alpha = 0.9) + 
    facet_wrap(~facet, scales = "free_x") +
    theme_hc() + scale_colour_hc(name = "") +
    theme(strip.text = element_text(hjust = 0), 
          strip.background = element_rect(fill = "grey90"),
          legend.position = "top",
          axis.ticks.y = element_blank()) + 
    xlab("Generation") + ylab("Frequency")
}

fig2 <- plot_run()
```

```{r fig.height=6, fig.width=10}
fig2
```
<br></br>
**Figure 2**: Three illustrative runs of the simulation, showing evolution in response to the introduction of 20 males carrying a _W_-shredder at Generation 50 (marked by the dotted line). In panel A, the driving _Z\*_ allele fixed very quickly, causing population extinction as the number of females dropped to zero. In panel B, the _Z\*_ allele spread up until the point that its fitness costs began to negate its transmission advantage, causing the population to halve in size but not to go extinct. In panel C, the _Z\*_ allele invaded, selecting for the resistance alleles _A_ and _Zr_, and causing the _Z\*_ allele to reverse course and go extinct. The population size _N_ is shown as a fraction of its maximum value of 10,000. Table S3 gives the parameter spaces used for these three runs. 

```{r echo=FALSE, results='hide'}
cairo_pdf(file = "figures/figure2.pdf", width = 10, height = 6); fig2; dev.off()
```



```{r}
tabl <- t(rbind(
  data.frame(Panel = "A", W_shredder %>% filter(id == "10024119427447")),
  data.frame(Panel = "B", W_shredder %>% filter(id == "1002596940286")),
  data.frame(Panel = "C", W_shredder %>% filter(id == "10024112958792"))) %>%
  select(release_strategy, W_shredding_rate, 
         Z_conversion_rate, Zr_creation_rate, Zr_mutation_rate,
         Wr_mutation_rate, cost_Zdrive_female, cost_Zdrive_male,
         male_migration_prob, female_migration_prob,
         migration_type, n_patches, softness, male_weighting,
         density_dependence_shape, max_fecundity, initial_A, initial_B))

nice_names2 <- data.frame(original = variable_parameters,
                          new = gsub("_", " ", variable_parameters),
                          stringsAsFactors = FALSE) %>%
  mutate(
    new = gsub("rel", "Rel", new),
    new = gsub("W shredding rate", "Gene drive in females (p_shred)", new),
    new = gsub("Z conversion rate", "Gene drive in males (p_conv)", new),
    new = gsub("Zr creation rate", "Rate of NHEJ (p_nhej)", new),
    new = gsub("Zr mutation rate", "Z-linked resistance (μ_Z)", new),
    new = gsub("Wr mutation rate", "W-linked resistance (μ_W)", new),
    new = gsub("cost Zdrive female", "Cost of Z* to females (c_f)", new),
    new = gsub("cost Zdrive male", "Cost of Z* to males (c_m)", new),
    new = gsub("male migration prob", "Male dispersal frequency", new),
    new = gsub("feMale dispersal frequency", "Female dispersal frequency (x_f)", new),
    new = gsub("Male dispersal frequency", "Male dispersal frequency (x_m)", new),
    new = gsub("migration type", "Dispersal type", new),
    new = gsub("n patches", "Number of patches (k)", new),
    new = gsub("softness", "Importance of local density (ψ)", new),
    new = gsub("male weighting", "Male effect on density (𝛿)", new),
    new = gsub("density dependence shape", "Shape of density dependence (α)", new),
    new = gsub("max fecundity", "Maximum fecundity (r)", new),
    new = gsub("initial A", "Intial freq. W-shredding resistance allele A", new),
    new = gsub("initial B", "Intial freq. gene conversion resistance allele B", new)) 

rownames(tabl) <- nice_names2$new[match(rownames(tabl), nice_names2$original)]
rownames(tabl) <- gsub("[.]", "", rownames(tabl))
colnames(tabl) <- c("Panel A", "Panel B", "Panel C")
```

**Table S3**: List of the parameter values used to generate the simulation runs shown in Figure 1.
<br></br>
```{r}
as.data.frame(tabl) %>% kable(format = "html", escape = FALSE) %>% kable_styling()
```



## Plot the effects of each parameter on extinction

```{r fig.width = 9, fig.height = 13}
# Find the proportion of runs that went extinct for each parameter value
get_percent_extinct <- function(dat, parameter){
  dat %>%
    group_by(!! sym(parameter)) %>%
    summarise(extinct = sum(went_extinct),
              not_extinct = n() - sum(went_extinct),
              prob = list(binom.test(extinct, extinct + not_extinct))) %>%
    rowwise() %>%
    mutate(response_variable = 100 * extinct/(extinct + not_extinct),
           lower_95_CI = 100 * prob$conf.int[1],
           upper_95_CI = 100 * prob$conf.int[2],
           parameter =  parameter) %>% rename(value = !! sym(parameter)) %>%
    select(parameter, value, everything()) %>% select(-prob) %>% ungroup()
} 

# Find the median time to extinction, among just those runs in which extinction occurred, for each parameter value
get_extinction_time <- function(dat, parameter){
  dat %>%
    filter(!is.na(generation_extinct)) %>% # only include runs where extinction occurred
    group_by(!! sym(parameter)) %>%
    summarise(response_variable = median(generation_extinct),
              lower_95_CI = quantile(generation_extinct, probs = 0.025),
              upper_95_CI = quantile(generation_extinct, probs = 0.975),
              n = n(),
              parameter =  parameter) %>% rename(value = !! sym(parameter)) %>%
    select(parameter, value, everything()) %>% ungroup()
} 

# To calculate separately for the 3 levels of Z_conversion_rate (male gene drive)
split_by_male_drive <- function(dat, fun){
  rbind(
    lapply(
      variable_parameters[variable_parameters != "Z_conversion_rate"], 
      get(fun), 
      dat = dat %>%
        filter(Z_conversion_rate == 0)) %>% do.call("rbind", .) %>% mutate(Z_conversion_rate = 0),
    lapply(
      variable_parameters[variable_parameters != "Z_conversion_rate"], 
      get(fun), 
      dat = dat %>%
        filter(Z_conversion_rate == 0.5)) %>% do.call("rbind", .) %>% mutate(Z_conversion_rate = 0.5),
    lapply(
      variable_parameters[variable_parameters != "Z_conversion_rate"], 
      get(fun), 
      dat = dat %>%
        filter(Z_conversion_rate == 0.95)) %>% do.call("rbind", .) %>% mutate(Z_conversion_rate = 0.95)
  ) %>% mutate(Z_conversion_rate = factor(Z_conversion_rate, levels = c(0, 0.5, 0.95)))
}

# % extinct, calculated separately for the three levels of Z_conversion_rate
W_shredder_split_by_male_drive_percent_extinct <- split_by_male_drive(W_shredder, "get_percent_extinct")
female_sterilising_split_by_male_drive_percent_extinct <- split_by_male_drive(female_sterilising, "get_percent_extinct") %>%
  filter(!(parameter %in% c("cost_Zdrive_female", "W_shredding_rate")))

# Time until extinction, calculated separately for the three levels of Z_conversion_rate
W_shredder_split_by_male_drive_time_to_extinct <- split_by_male_drive(W_shredder, "get_extinction_time")
female_sterilising_split_by_male_drive_time_to_extinct <- split_by_male_drive(female_sterilising, "get_extinction_time") %>%
  filter(!(parameter %in% c("cost_Zdrive_female", "W_shredding_rate")))

# Function to make Figure 3 (and similar supplementary figures)
make_multipanel_figure <- function(dat, ylab){
  parameter_importance <- dat %>%
    mutate(parameter = nice_names$new[match(parameter, nice_names$original)]) %>%
    group_by(parameter) %>% summarise(range = max(response_variable) - min(response_variable)) %>%
    arrange(-range) %>% pull(parameter)
  
  levels <- levels(factor(dat$value))
  levels <- c(levels[!(levels %in% c("all_patches", 100))], 100, "all_patches")
  levels <- replace(levels, levels == "all_patches", "Release in\nall patches")
  levels <- replace(levels, levels == "one_patch", "Release in\na single patch")
  
  pd <- position_dodge(0.2)

  dat %>% 
    mutate(parameter = nice_names$new[match(parameter, nice_names$original)],
           parameter = factor(parameter, parameter_importance),
           value = replace(value, value == "all_patches", "Release in\nall patches"),
           value = replace(value, value == "one_patch", "Release in\na single patch"),
           value = factor(value, levels)) %>%
    ggplot(aes(value, response_variable, colour = Z_conversion_rate)) + 
    geom_errorbar(aes(ymin = lower_95_CI, ymax = upper_95_CI), width = 0, position = pd) +
    geom_point(position = pd) + 
    scale_colour_brewer(palette = "Set2", name = "Gene drive in males\n(p_conv)") + 
    xlab("Parameter value") +
    ylab(ylab) + 
    theme_bw() + 
    theme(strip.background = element_blank(),
          panel.grid.major.x = element_blank(),
          panel.border = element_rect(fill = NA, size = 1))
  
}

fig3 <- make_multipanel_figure(W_shredder_split_by_male_drive_percent_extinct, 
                      ylab = "% simulation runs resulting in extinction (\u00B1 95% CIs)")
```


### % extinction plot for a W-shredder (Figure 3)
```{r fig.width = 9, fig.height = 12}
cairo_pdf(file = "figures/figure3.pdf", width = 9, height = 12)
fig3 + facet_wrap(~parameter, scales = "free", labeller = label_parsed, ncol = 4) + theme(legend.position = c(.9, .1))
dev.off()

fig3 + facet_wrap(~parameter, scales = "free", labeller = label_parsed, ncol = 3) 
```
<br></br>
**Figure 3:** The percentage of simulations of a _W_-shredder that ended in extinction, for all runs with a particular value (shown on the $x$-axis) for a given parameter (shown in the panels). For example, in all the thousands of runs for which I assumed $p_{shred} = 0.5$, there were no extinctions, while among the runs where $p_{shred} = 1$, over 60% resulted in extinction. The panels are ordered by the range of the y-axis, which highlights the relative importance of the variables for the probability of extinction. Figure S1 shows the median time-to-extinction for each parameter value, and Figures S3-S4 give similar plots for simulations that considered a female-sterilising _Z_-linked gene drive instead of a _W_-shredder.

### Time-to-extinction plot for a W-shredder (Figure S1)

```{r fig.height = 12, fig.width = 10}
S1_fig <- make_multipanel_figure(filter(W_shredder_split_by_male_drive_time_to_extinct, n >= 40), 
                        ylab = "Median generations until extinction (\u00B1 95% quantiles)") + 
  facet_wrap(~parameter, scales = "free", labeller = label_parsed, ncol = 3) + 
  coord_cartesian(ylim = c(0, 150))

ggsave(S1_fig, filename = "figures/S1_fig.pdf", height = 12, width = 19.5)
saveRDS(S1_fig, file = "figures/S1_fig.rds")
S1_fig
```
<br></br>
**Figure S1:** Similar plot to Figure 3, except that the y-axis shows the median number of generations until the _W_-shredder caused extinction, among just the subset of simulations in which extinction actually occurred. A median was only calculated if at least 40 simulation runs reached extinction.

### % extinction plot for a female-sterilising Z-drive (Figure S3)

```{r fig.width = 10, fig.height = 9.6}
S3_fig <- make_multipanel_figure(female_sterilising_split_by_male_drive_percent_extinct,
                        ylab = "% simulation runs resulting in extinction (\u00B1 95% CIs)") + 
  facet_wrap(~parameter, scales = "free", labeller = label_parsed, ncol = 4)

ggsave(S3_fig, filename = "figures/S3_fig.pdf", height = 9.6, width = 10)
saveRDS(S3_fig, file = "figures/S3_fig.rds")
S3_fig
```
<br></br>
**Figure S3**: Analagous information to Figure 3, but showing the results for a female-sterilising _Z\*_ allele instead of a _W_-shredder.


### Time-to-extinction plot for a female-sterilising Z-drive (Figure S4)
```{r fig.height = 12, fig.width = 10}
S4_fig <- make_multipanel_figure(filter(female_sterilising_split_by_male_drive_time_to_extinct, n >= 40), 
                        ylab = "Median generations until extinction (\u00B1 95% quantiles)") + 
  facet_wrap(~parameter, scales = "free", labeller = label_parsed, ncol = 3) + 
  coord_cartesian(ylim = c(0, 100))

ggsave(S4_fig, filename = "figures/S4_fig.pdf", height = 12, width = 10)
saveRDS(S4_fig, file = "figures/S4_fig.rds")
S4_fig
```
<br></br>
**Figure S4**: Analagous information to Figure S1, but showing the results for a female-sterilising _Z\*_ allele instead of a _W_-shredder. Note that a median was only calculated if at least 40 simulation runs reached extinction, and extinction only occurred when gene drive in males was strong.


## Finding variables with interacting effects on the extinction probability

### Testing for interactions using GLM 
Run a binomial generalized linear model (GLM), with all of the variable model parameters and all their 2-way interactions as predictors. The response variable is extinction, coded as a zero or one. Since the number of data points is very large (millions), I use the `biglm` package ("big GLM"). I run a model separately on all the runs that considered the evolution of a _W_-shredder, and all the runs that considered a female-sterilising _Z\*_.  

```{r big_glm}
run_big_glm <- function(dat, sterilising = FALSE){
  
  if(sterilising){
    variable_parameters <- variable_parameters[!(variable_parameters %in% c("W_shredding_rate", "cost_Zdrive_female"))] 
  }
  
  formula <- paste("went_extinct ~ (",
                   paste0(variable_parameters, collapse = " + "), ")^2",
                   sep = "")
  
  if(!sterilising) formula <- paste(formula, "- W_shredding_rate:cost_Zdrive_female")
  
  bigglm(as.formula(formula), 
         data = dat %>% mutate_if(is.numeric, function(x) as.numeric(scale(x))), 
         chunksize = 10000)
}

# Helper to get the fixed effects table out of the biglm object (no 'coefficients' slot, apparently)
tidy_model <- function(model){
  
  out <- capture.output(model %>% summary)[-(1:5)] %>% lapply(function(x) {
    x <- strsplit(x, split = " ")[[1]]
    x[x != ""]
    }) %>% do.call("rbind", .) %>% as.data.frame(stringsAsFactors = FALSE) %>% as_tibble() %>% 
    mutate(V1 = as.factor(gsub(":", " x ", V1))) %>% mutate_if(is.character, as.numeric) 
  
  names(out) <- c("Parameter", "Estimate", "Lower_95_CI", "Upper_95_CI", "SE", "p")
  out %>% arrange(-abs(Estimate))
}

if(!file.exists("data/W_shredder_model.rds")) {
  W_shredder_model <- run_big_glm(W_shredder)
  saveRDS(tidy_model(W_shredder_model), "data/W_shredder_model.rds")
  female_sterilising_model <- run_big_glm(female_sterilising, sterilising = TRUE)
  saveRDS(tidy_model(female_sterilising_model), "data/female_sterilising_model.rds")
} else {
  W_shredder_model <- readRDS("data/W_shredder_model.rds")
  female_sterilising_model <- readRDS("data/female_sterilising_model.rds")
}
```


### Finding the top-ranked effects in the GLM

```{r fig.height = 9}
importance_plot <- function(dat, n = "all"){
  
  if(n != "all"){
    dat <- dat %>%
      mutate(row=1:n()) %>% filter(row %in% 1:n)
  } 
  
  for(i in 1:nrow(dat)){
    if(dat$Estimate[i] < 0){
      lower <- dat$Lower_95_CI[i]
      dat$Lower_95_CI[i] <- dat$Upper_95_CI[i]
      dat$Upper_95_CI[i] <- lower
    }
  }
  
  dat <- dat %>%
    mutate(Parameter = gsub("one_patch", "", as.character(Parameter)),
           Parameter = gsub("Local", "", Parameter),
           Parameter = factor(Parameter, rev(Parameter)),
           tick_label = nice_names$new[match(Parameter, nice_names$original)]) 
  
  dat %>%
    ggplot(aes(Parameter, abs(Estimate))) + 
    geom_bar(stat = "identity", fill = "tomato") +
    geom_errorbar(aes(ymin = abs(Lower_95_CI), ymax = abs(Upper_95_CI)), width = 0) + 
    coord_flip() + 
    scale_y_continuous(expand = c(0, 0)) + 
    scale_x_discrete(labels = parse(text = rev(dat$tick_label))) + 
    xlab("MPredictor variable) + ylab("Absolute effect size (\u00B1 95% CIs)") + 
    theme_classic() + 
    theme(axis.ticks.y = element_blank())
}

W_shred        <- importance_plot(W_shredder_model, n = 25)
female_sterile <- importance_plot(female_sterilising_model, n = 25)
```

#### For a _W_-shredder (Figure 4)

```{r fig.height = 8, fig.width = 8}
W_shred
```
<br></br>
**Figure 4**: Relative parameter importance in the simulations of _W_-shredders, for the top 25 most important main effects or two-way interactions (from a binomial GLM that included all the main effects and all their two-way interactions). Each predictor variable was scaled before running the model, meaning that the absolute effect size indicates how important each parameter is to the extinction probability, given the range of values plotted in Figure 3. Figure S5 gives a similar plot for simulations of a female-sterilising $Z^*$ allele.

```{r echo=FALSE, results='hide'}
ggsave(W_shred, file = "figures/figure4.pdf", height = 8, width = 8)
```

#### For a female-sterilising drive

```{r fig.height = 8, fig.width = 8}
female_sterile
```
<br></br>
**Figure S5**: Relative parameter importance in the simulations of _Z_-linked female-sterilising gene drives, for the top 25 most important main effects or two-way interactions (from a binomial GLM that included all the main effects and all their two-way interactions). Each predictor variable was scaled before running the model, meaning that the absolute effect size indicates how important each parameter is to the extinction probability, given the range of values plotted in Figure S3.

```{r echo=FALSE, results='hide'}
ggsave(female_sterile, file = "figures/S5_fig.pdf", height = 8, width = 8)
saveRDS(female_sterile, file = "figures/S5_fig.rds")
```

### Plot showing interacting effects on extinction probability

```{r twowayplot, fig.width = 15, fig.height = 15}
get_percent_extinct_double <- function(dat, p1, p2){
  output <-  dat %>%
    group_by(!! sym(p1), !! sym(p2)) %>%
    summarise(extinct = sum(went_extinct),
              not_extinct = n() - sum(went_extinct),
              prob = list(binom.test(extinct, extinct + not_extinct))) %>%
    rowwise() %>%
    mutate(percent_extinct = 100 * extinct / (extinct + not_extinct),
           lower_95_CI = 100 * prob$conf.int[1],
           upper_95_CI = 100 * prob$conf.int[2],
           parameter_1 =  p1, 
           parameter_2 =  p2) 
  names(output)[1:2] <- paste("value", 1:2, sep = "_")
  output %>%
    mutate(value_1 = as.character(value_1), value_2 = as.character(value_2)) %>%
    select(parameter_1, parameter_2, value_1, value_2, everything()) %>% 
    select(-prob) %>% ungroup()
} 

top_12_shredder    <- (W_shredder_model %>% mutate(Parameter = as.character(Parameter)) %>% filter(grepl(" x ", Parameter)) %>% pull(Parameter))[1:12]
top_12_sterilising <- (female_sterilising_model %>% mutate(Parameter = as.character(Parameter)) %>% filter(grepl(" x ", Parameter)) %>% pull(Parameter))[1:12]

two_way_W_shredder <- map2_df(str_split(top_12_shredder, " x ", simplify = TRUE)[, 1], 
                              str_split(top_12_shredder, " x ", simplify = TRUE)[, 2], 
                              get_percent_extinct_double, dat = W_shredder)

two_way_sterilising <- map2_df(str_split(top_12_sterilising, " x ", simplify = TRUE)[, 1], 
                               str_split(top_12_sterilising, " x ", simplify = TRUE)[, 2], 
                               get_percent_extinct_double, dat = W_shredder)



two_way_plot <- function(two_way_data, pairwise_combos){
  
  two_way_data <- two_way_data %>%
    mutate(pasted = paste(parameter_1, parameter_2, sep = " x "),
           parameter_1 = nice_names$new[match(parameter_1, nice_names$original)],
           parameter_2 = nice_names$new[match(parameter_2, nice_names$original)]) %>%
    filter(pasted %in% pairwise_combos) 
  z_range <- range(two_way_data$percent_extinct) * 1.1
  
  make_one <- function(df){
    
    df$value_1 <- factor(df$value_1, unique(df$value_1))
    df$value_2 <- factor(df$value_2, unique(df$value_2))
    
    ggplot(df, aes(value_1, value_2, fill = percent_extinct)) + 
      geom_tile(colour = "black", linetype = 3) + 
      scale_fill_distiller(palette = "PuBuGn", direction = 1, limits = z_range, name = "% extinct") + 
      theme_bw() + 
      theme() + 
      scale_x_discrete(expand = c(0,0), name = parse(text = df$parameter_1[1])) + 
      scale_y_discrete(expand = c(0,0), name = parse(text = df$parameter_2[1])) 
      
  }
  
  grid_arrange_shared_legend <- function(...) {
    plots <- list(...)
    g <- ggplotGrob(plots[[1]] + theme(legend.position="bottom"))$grobs
    legend <- g[[which(sapply(g, function(x) x$name) == "guide-box")]]
    lheight <- sum(legend$height)
    plots <- lapply(plots, function(x) ggplotGrob(x + theme(legend.position = "none")))
    
    p <- gtable_cbind(gtable_rbind(plots[[1]], plots[[4]], plots[[7]]),
                      gtable_rbind(plots[[2]], plots[[5]], plots[[8]]),
                      gtable_rbind(plots[[3]], plots[[6]], plots[[9]]))
    
    grid.arrange(p, legend, 
                 nrow = 2, 
                 heights = unit.c(unit(1, "npc") - lheight, lheight))
  }

  dat_list <- vector(mode = "list", length = length(pairwise_combos))
  for(i in 1:length(pairwise_combos)){
    dat_list[[i]] <- two_way_data %>% filter(pasted == pairwise_combos[i])
  }
  
  lapply(dat_list, make_one) %>% do.call("grid_arrange_shared_legend", .)
}

two_way_plot_shredder <- two_way_plot(two_way_W_shredder, top_12_shredder)
two_way_plot_sterilising <- two_way_plot(two_way_sterilising, top_12_sterilising)
saveRDS(two_way_plot_shredder, file = "figures/S2_fig.rds")
saveRDS(two_way_plot_sterilising, file = "figures/S6_fig.rds")
grid.draw(two_way_plot_shredder)
```
<br></br>
**Figure S2:** Heatmap illustrating the twelve strongest two-way interactions for simulations of a _W_-shredder, as determined by the effect sizes from the GLM plotted in Figure 4. 


```{r fig.width = 15, fig.height = 15}
grid.draw(two_way_plot_sterilising)
```
**Figure S6:** Heatmap illustrating the twelve strongest two-way interactions for simulations of a female-sterilising gene drive, as determined by the effect sizes from the GLM plotted in Figure S5. 



```{r echo=FALSE, results='hide'}
cairo_pdf(file = "figures/S2_fig.pdf", width = 15, height = 15); grid.draw(two_way_plot_shredder); dev.off()
cairo_pdf(file = "figures/S6_fig.pdf", width = 15, height = 15); grid.draw(two_way_plot_sterilising); dev.off()
```


