---
title: "Plotting the simulation results"
output: workflowr::wflow_html
editor_options:
  chunk_output_type: console
---

```{r setup, echo=FALSE, results='hide'}
source_rmd <- function(file, local = FALSE){
  options(knitr.duplicate.label = "allow")
  tempR <- tempfile(tmpdir = ".", fileext = ".R")
  on.exit(unlink(tempR))
  print(file.exists("model_functions.Rmd"))
  knitr::purl(file, output = tempR, quiet = TRUE)
  source(tempR, local = globalenv())
}
source_rmd("analysis/model_functions.Rmd")
```

## Load packages and results

```{r results='hide', message=FALSE, warning=FALSE}
packages <- c("dplyr", "purrr", "ggplot2", "reshape2", "Cairo", "knitr",  
              "latex2exp", "pander", "grid", "gridExtra", "ggthemes",
              "readr", "tibble", "biglm", "kableExtra")
shh <- suppressMessages(lapply(packages, library, character.only = TRUE, quietly = TRUE))


if(!file.exists("data/all_results.rds")){
  results <- lapply(list.files(path = "data", pattern = "results_", full.names = TRUE), readRDS) %>% 
    do.call("rbind", .) %>%
    filter(cost_A == 0 & cost_B == 0)
  saveRDS(results, file = "data/all_results.rds")
} else results <- read_rds("data/all_results.rds")

results$went_extinct <- ifelse(results$outcome == "Population extinct", 1, 0)

results$migration_type[results$migration_type == "local"] <- "Local"
results$migration_type[results$migration_type == "global"] <- "Global"

find_variable_parameters <- function(dat){
  dat %>% 
    select(-id, -realisations, -generation_extinct, -generation_Zd_extinct, 
           -generation_W_extinct, -generation_Zd_fixed, -outcome, -went_extinct, 
           -mating_table, -initial_Wr, -initial_Zr) %>%
    sapply(function(x) length(unique(x))) %>% 
    keep(~.x > 1) %>% names()
}

variable_parameters <- find_variable_parameters(results)
combinations <- apply(combn(variable_parameters, 2), 2, paste0, collapse = " x ")

# Make a data frame to conver R-friendly names to figure-friendly names
nice_names <- data.frame(original = c(variable_parameters, combinations),
                         new = gsub("_", " ", c(variable_parameters, combinations)),
                         stringsAsFactors = FALSE) %>%
  mutate(
    new = gsub("rel", "Rel", new),
    new = gsub("W shredding rate", "Gene drive in females ($p_{shred}$)", new),
    new = gsub("Z conversion rate", "Gene drive in males ($p_{conv}$)", new),
    new = gsub("Zr creation rate", "Rate of NHEJ ($p_{nhej}$)", new),
    new = gsub("Zr mutation rate", "Z-linked resistance ($\\mu{_Z}$)", new),
    new = gsub("Wr mutation rate", "W-linked resistance ($\\mu{_W}$)", new),
    new = gsub("cost Zdrive female", "Cost of Z* to females ($c_f$)", new),
    new = gsub("cost Zdrive male", "Cost of Z* to males ($c_m$)", new),
    new = gsub("male migration prob", "Male dispersal frequency", new),
    new = gsub("feMale dispersal frequency", "Female dispersal frequency (xf)", new),
    new = gsub("Male dispersal frequency", "Male dispersal frequency (xm)", new),
    new = gsub("xm", "$x_m$", new), 
    new = gsub("xf", "$x_f$", new),
    new = gsub("migration type", "Dispersal type", new),
    new = gsub("n patches", "Number of patches ($k$)", new),
    new = gsub("softness", "Importance of local density ($\\psi$)", new),
    new = gsub("male weighting", "Male effect on density ($\\delta$)", new),
    new = gsub("density dependence shape", "Shape of density dependence ($\\alpha$)", new),
    new = gsub("max fecundity", "Maximum fecundity ($r$)", new),
    new = gsub("initial A", "Intial freq. W-shredding resistance allele A", new),
    new = gsub("initial B", "Intial freq. gene conversion resistance allele B", new),
    new = as.character(TeX(new))) %>% 
  mutate(new = gsub("mu", "\\mu", new))

# Sort the results between those that model a W-shredder, versus those that model a Z* that sterilisies females
W_shredder <- results %>% filter(cost_Zdrive_female != 1)
female_sterilising <- results %>% filter(cost_Zdrive_female == 1)
rm(results)
variable_parameters_W_shredder <- find_variable_parameters(W_shredder)
variable_parameters_female_sterilising <- find_variable_parameters(female_sterilising)
```

## Table showing the frequencies of each possible outcome

**Table S1**: The number and percentage of simulation runs that ended with the five possible outcomes, for the subset of simulation runs focusing on a _W_-shredder gene drive.
<br></br>
```{r}
make_tally_table <- function(dat){
  outcomes <- dat$outcome %>% table() %>% melt() %>% arrange(-value) %>% mutate(p = round(100 * value / sum(value), 1))
  names(outcomes) <- c("Outcome", "Number of simulations", "%")
  outcomes %>% mutate(Outcome = gsub("without", "without causing", Outcome),
                      Outcome = gsub("n extinct", "n went extinct", Outcome),
                      Outcome = gsub("d extinct", "d went extinct", Outcome)) %>% pander()
}

make_tally_table(W_shredder)
```

**Table S2**: The number and percentage of simulation runs that ended with the five possible outcomes, for the subset of simulation runs focusing on a female-sterilising _Z_-linked gene drive.
<br></br>
```{r}
make_tally_table(female_sterilising)
```

## Make Figure 1

```{r}
a <- read_rds("data/allele_freqs_1.rds")

plot_run <- function(){
  
  get_data <- function(model_id, allele_freqs_list, result_df, label){
    
    df <- pluck_allele_freqs(model_id, allele_freqs_list) 
    
    alleles_to_plot <- group_by(df, allele) %>% 
      summarise(uniques = length(unique(frequency))) %>% 
      filter(uniques > 1) %>% # Don't plot alleles that stay at 0 whole time
      pull(allele)
    
    alleles_to_plot <- c(alleles_to_plot, "N")
    df <- df[df$allele %in% alleles_to_plot, ]
    df$frequency[df$allele == "N"] <- df$frequency[df$allele == "N"] /
      max(df$frequency[df$allele == "N"])
     
    paras <- result_df %>% mutate(id = as.character(id)) %>% filter(id == model_id)
    last_generation <- tail(df,12) %>% select(allele, frequency)
    print(paras); print(last_generation)

    df %>% mutate(facet = label,
                  allele = replace(allele, allele == "Zd", "Z*"),
                  allele = replace(allele, allele == "females", "Females"))
  }
  
  rbind(
    get_data("10024119427447", a, W_shredder, "A. Population extinction"),
    get_data("1002596940286", a, W_shredder, "B. Failure to cause extinction"),
    get_data("10024112958792", a, W_shredder, "C. Resistance prevents extinction")) %>%
    ggplot(aes(x = generation, y = frequency, colour = allele, group = allele)) + 
    geom_vline(xintercept = 50, linetype = 2, colour = "grey10", size = 0.9) + 
    geom_line(size = 0.9, alpha = 0.9) + 
    facet_wrap(~facet, scales = "free_x") +
    theme_hc() + scale_colour_hc(name = "") +
    theme(strip.text = element_text(hjust = 0), 
          strip.background = element_rect(fill = "grey90"),
          legend.position = "top",
          axis.ticks.y = element_blank()) + 
    xlab("Generation") + ylab("Frequency")
}

fig1 <- plot_run()
rm(a)
fig1
```
<br></br>
**Figure 1**: Three illustrative runs of the simulation, showing evolution in response to the introduction of XX males carrying a _W_-shredder at Generation 50 (marked by the dashed line). In panel A, the driving _Z\*_ allele fixed very quickly, causing population extinction as the number of females dropped to zero. In panel B, the _Z\*_ allele spread up until the point that its fitness costs began to negate its transmission advantage, causing the population to halve in size but not to go extinct. In panel C, the _Z\*_ allele invaded, selecting for the resistance alleles _A_ and _Zr_ allele, and causing the _Z\*_ allele to reverse course and go extinct. The population size _N_ is shown as a fraction of its maximum value of 10,000. Table S3 gives the parameter spaces used for these three runs. 

```{r echo=FALSE, results='hide'}
cairo_pdf(file = "figures/figure1.pdf", width = 10, height = 10); fig1; dev.off()
```

<br></br>
```{r}
tabl <- t(rbind(
  data.frame(Panel = "A", W_shredder %>% filter(id == "10024119427447")),
  data.frame(Panel = "B", W_shredder %>% filter(id == "1002596940286")),
  data.frame(Panel = "C", W_shredder %>% filter(id == "10024112958792"))) %>%
  select(release_strategy, W_shredding_rate, 
         Z_conversion_rate, Zr_creation_rate, Zr_mutation_rate,
         Wr_mutation_rate, cost_Zdrive_female, cost_Zdrive_male,
         male_migration_prob, female_migration_prob,
         migration_type, n_patches, softness, male_weighting,
         density_dependence_shape, max_fecundity, initial_A, initial_B))

nice_names2 <- data.frame(original = variable_parameters,
                          new = gsub("_", " ", variable_parameters),
                          stringsAsFactors = FALSE) %>%
  mutate(
    new = gsub("rel", "Rel", new),
    new = gsub("W shredding rate", "Gene drive in females (p_shred)", new),
    new = gsub("Z conversion rate", "Gene drive in males (p_conv)", new),
    new = gsub("Zr creation rate", "Rate of NHEJ (p_nhej)", new),
    new = gsub("Zr mutation rate", "Z-linked resistance (μ_Z)", new),
    new = gsub("Wr mutation rate", "W-linked resistance (μ_W)", new),
    new = gsub("cost Zdrive female", "Cost of Z* to females (c_f)", new),
    new = gsub("cost Zdrive male", "Cost of Z* to males (c_m)", new),
    new = gsub("male migration prob", "Male dispersal frequency", new),
    new = gsub("feMale dispersal frequency", "Female dispersal frequency (x_f)", new),
    new = gsub("Male dispersal frequency", "Male dispersal frequency (x_m)", new),
    new = gsub("migration type", "Dispersal type", new),
    new = gsub("n patches", "Number of patches (k)", new),
    new = gsub("softness", "Importance of local density (ψ)", new),
    new = gsub("male weighting", "Male effect on density (𝛿)", new),
    new = gsub("density dependence shape", "Shape of density dependence (α)", new),
    new = gsub("max fecundity", "Maximum fecundity (r)", new),
    new = gsub("initial A", "Intial freq. W-shredding resistance allele A", new),
    new = gsub("initial B", "Intial freq. gene conversion resistance allele B", new)) 



rownames(tabl) <- nice_names2$new[match(rownames(tabl), nice_names2$original)]
rownames(tabl) <- gsub("[.]", "", rownames(tabl))
colnames(tabl) <- c("Panel A", "Panel B", "Panel C")
```

**Table S3**: List of the parameter values used to generate the simulation runs shown in Figure 1.
```{r}
as.data.frame(tabl) %>% kable(format = "html", escape = FALSE) %>% kable_styling()
```



## Make Figure 2

```{r fig.width = 9, fig.height = 13}
# Find the proportion of runs that went extinct for each parameter value
get_percent_extinct <- function(dat, parameter){
  dat %>%
    group_by(!! sym(parameter)) %>%
    summarise(extinct = sum(went_extinct),
              not_extinct = n() - sum(went_extinct),
              prob = list(binom.test(extinct, extinct + not_extinct))) %>%
    rowwise() %>%
    mutate(percent_extinct = 100 * extinct/(extinct + not_extinct),
           lower_95_CI = 100 * prob$conf.int[1],
           upper_95_CI = 100 * prob$conf.int[2],
           parameter =  parameter) %>% rename(value = !! sym(parameter)) %>%
    select(parameter, value, everything()) %>% select(-prob) %>% ungroup()
} 

W_shredder_percent_extinct <- lapply(variable_parameters, 
                                     get_percent_extinct, 
                                     dat = W_shredder) %>% do.call("rbind", .)

female_sterilising_percent_extinct <- lapply(variable_parameters, 
                                             get_percent_extinct, 
                                             dat = female_sterilising) %>% do.call("rbind", .)

# Graph these data
make_figure_2 <- function(percent_extinct_data){
  parameter_importance <- percent_extinct_data %>%
    mutate(parameter = nice_names$new[match(parameter, nice_names$original)]) %>%
    group_by(parameter) %>% summarise(range = max(percent_extinct) - min(percent_extinct)) %>%
    arrange(-range) %>% pull(parameter)
  
  levels <- levels(factor(percent_extinct_data$value))
  levels <- c(levels[!(levels %in% c("all_patches", 100))], 100, "all_patches")
  levels <- replace(levels, levels == "all_patches", "Release in\nall patches")
  levels <- replace(levels, levels == "one_patch", "Release in\na single patch")
  
  percent_extinct_data %>% 
    mutate(parameter = nice_names$new[match(parameter, nice_names$original)],
           parameter = factor(parameter, parameter_importance),
           value = replace(value, value == "all_patches", "Release in\nall patches"),
           value = replace(value, value == "one_patch", "Release in\na single patch"),
           value = factor(value, levels)) %>%
    ggplot(aes(value, percent_extinct)) + 
    geom_errorbar(aes(ymin=lower_95_CI, ymax = upper_95_CI), width = 0) +
    geom_point() + 
    xlab("Parameter value") +
    ylab("% simulation runs resulting in extinction (\u00B1 95% CIs)") + 
    theme_bw() + 
    theme(strip.background = element_blank(),
          panel.grid.major.x = element_blank())
  
}

fig2 <- make_figure_2(W_shredder_percent_extinct) 
fig2 + facet_wrap(~parameter, scales = "free", labeller = label_parsed, ncol = 3) 
```
<br></br>
**Figure 2:** The percentage of simulations of a _W_-shredder that ended in extinction, for all runs with a particular value (shown on the x-axis) for a given parameter (shown in the panels). For example, in all the thousands of runs for which I assumed $p_{shred} = 0.5$, there were no extinctions, while among the runs where $p_{shred} = 1$, over 60% resulted in extinction. The panels are ordered by the range of the y-axis, which highlights the relative importance of the variables for the probability of extinction. Figure S1 gives a similar plot for simulations of a female-sterilising _Z\*_ allele.

```{r echo=FALSE, results='hide'}
cairo_pdf(file = "figures/figure2.pdf", width = 9, height = 12)
fig2 + facet_wrap(~parameter, scales = "free", labeller = label_parsed, ncol = 4)
dev.off()
```


## Make Figure S1

```{r fig.width = 9, fig.height = 13}
figS1 <- make_figure_2(female_sterilising_percent_extinct)
cairo_pdf(file = "figures/figureS1.pdf", width = 9, height = 12)
figS1 + facet_wrap(~parameter, scales = "free", labeller = label_parsed, ncol = 4)
dev.off()

figS1 + facet_wrap(~parameter, scales = "free", labeller = label_parsed, ncol = 3) 
```
<br></br>
**Figure S1:** Analgous information to Figure 2, but showing the results for a female-sterilising _Z\*_ allele instead of a _W_-shredder.


## Finding variables with interacting effects on the extinction probability


### Testing for interactions using GLM 
Run a binomial generalized linear model (GLM), with all of the variable model parameters and all their 2-way interactions as predictors. The response variable is extinction, coded as a zero or one. Since the number of data points is very large (millions), I use the `biglm` package ("big GLM"). I run a model separately on all the runs that considered the evolution of a _W_-shredder (i.e. females are viable, but produce mostly sons), and the run that considered a female-sterilising _Z\*_.  

```{r}
run_big_glm <- function(dat, sterilising = FALSE){
  
  if(sterilising){
    variable_parameters <- variable_parameters[!(variable_parameters %in% c("W_shredding_rate", "cost_Zdrive_female"))] 
  }
  
  formula <- paste("went_extinct ~ (",
                   paste0(variable_parameters, collapse = " + "), ")^2",
                   sep = "")
  
  if(!sterilising) formula <- paste(formula, "- W_shredding_rate:cost_Zdrive_female")
  
  my_scale <- function(x) as.numeric(scale(x))
  
  glm_model <- bigglm(as.formula(formula), data = dat %>% mutate_if(is.numeric, my_scale), chunksize = 10000)
  return(glm_model)
  
  output <- data.frame(Parameter = rownames(summary(glm_model)[[2]]), summary(glm_model)[[2]]) %>% 
    as_tibble() %>% arrange(-abs(Coef)) %>%
    mutate(Parameter = gsub(":", " x ", Parameter),
           Parameter = factor(Parameter, rev(Parameter)),
           log10_p = -log10(p + 1e-300)) 
  
  names(output)[2:4] <- c("Estimate", "Lower_95_CI", "Upper_95_CI")
  output
}

if(!file.exists("data/W_shredder_model.rds")) {
  W_shredder_model <- run_big_glm(W_shredder)
  saveRDS(W_shredder_model, "data/W_shredder_model.rds")
  female_sterilising_model <- run_big_glm(female_sterilising, sterilising = TRUE)
  saveRDS(female_sterilising_model, "data/female_sterilising_model.rds")
} else {
  W_shredder_model <- readRDS("data/W_shredder_model.rds")
  female_sterilising_model <- readRDS("data/female_sterilising_model.rds")
}
```


### Finding the top-ranked effects in the GLM

```{r fig.height = 9}
importance_plot <- function(dat, n = "all"){
  if(n != "all") dat <- dat %>%
      mutate(row=1:n()) %>% filter(row %in% 1:n)
  
  for(i in 1:nrow(dat)){
    if(dat$Estimate[i] < 0){
      lower <- dat$Lower_95_CI[i]
      dat$Lower_95_CI[i] <- dat$Upper_95_CI[i]
      dat$Upper_95_CI[i] <- lower
    }
  }
  
  dat <- dat %>%
    mutate(Parameter = gsub("one_patch", "", as.character(Parameter)),
           Parameter = gsub("Local", "", Parameter),
           Parameter = factor(Parameter, rev(Parameter)),
           tick_label = nice_names$new[match(Parameter, nice_names$original)]) 
  
  dat %>%
    ggplot(aes(Parameter, abs(Estimate))) + 
    geom_bar(stat = "identity", fill = "tomato") +
    geom_errorbar(aes(ymin = abs(Lower_95_CI), ymax = abs(Upper_95_CI)), width = 0) + 
    coord_flip() + 
    scale_y_continuous(expand = c(0, 0)) + 
    scale_x_discrete(labels = parse(text = rev(dat$tick_label))) + 
    xlab("Model parameter") + ylab("Absolute effect size")

}
importance_plot(W_shredder_model, n = 25)
```
<br></br>
**Figure SX:**

```{r fig.height = 9}
importance_plot(female_sterilising_model, n = 25)
```
<br></br>
**Figure SX:**

### Plot showing interacting effects on extinction probability

```{r fig.width = 13, fig.height = 13}
get_percent_extinct_double <- function(dat, p1, p2){
  output <-  dat %>%
    group_by(!! sym(p1), !! sym(p2)) %>%
    summarise(extinct = sum(went_extinct),
              not_extinct = n() - sum(went_extinct),
              prob = list(binom.test(extinct, extinct + not_extinct))) %>%
    rowwise() %>%
    mutate(percent_extinct = 100 * extinct / (extinct + not_extinct),
           lower_95_CI = 100 * prob$conf.int[1],
           upper_95_CI = 100 * prob$conf.int[2],
           parameter_1 =  p1, 
           parameter_2 =  p2) 
  names(output)[1:2] <- paste("value", 1:2, sep = "_")
  output %>%
    mutate(value_1 = as.character(value_1), value_2 = as.character(value_2)) %>%
    select(parameter_1, parameter_2, value_1, value_2, everything()) %>% 
    select(-prob) %>% ungroup()
} 

# combinations <- combn(variable_parameters, 2)
# two_way_W_shredder <- map2_df(combinations[1, ], 
#                               combinations[2, ], 
#                               get_percent_extinct_double, dat = W_shredder)

interesting <- c("W_shredding_rate~initial_A", "Z_conversion_rate~cost_Zdrive_female", "W_shredding_rate~Wr_mutation_rate",
                 "W_shredding_rate~Z_conversion_rate", "Zr_creation_rate~cost_Zdrive_female", "W_shredding_rate~density_dependence_shape",
                 "cost_Zdrive_female~n_patches", "W_shredding_rate~max_fecundity", "density_dependence_shape~max_fecundity")
two_way_W_shredder <- map2_df(str_split(interesting, "~", simplify = TRUE)[, 1], 
                              str_split(interesting, "~", simplify = TRUE)[, 2], 
                              get_percent_extinct_double, dat = W_shredder)


two_way_plot <- function(two_way_data, pairwise_combos){
  
  two_way_data <- two_way_data %>%
    mutate(pasted = paste(parameter_1, parameter_2, sep="~"),
           parameter_1 = nice_names$new[match(parameter_1, nice_names$original)],
           parameter_2 = nice_names$new[match(parameter_2, nice_names$original)]) %>%
    filter(pasted %in% pairwise_combos) 
  z_range <- range(two_way_data$percent_extinct) * 1.1
  
  make_one <- function(df){
    
    ggplot(df, aes(value_1, value_2, fill = percent_extinct)) + 
      geom_tile(colour = "black", size = 1) + 
      scale_fill_distiller(palette = "PuBuGn", direction = 1, limits = z_range, name = "% extinct") + 
      theme_minimal() + 
      theme(panel.border = element_rect(fill = NA, colour = "black", size = 1)) + 
      scale_x_discrete(expand = c(0,0), name = parse(text = df$parameter_1[1])) + 
      scale_y_discrete(expand = c(0,0), name = parse(text = df$parameter_2[1])) 
      
  }
  
  grid_arrange_shared_legend <- function(...) {
    plots <- list(...)
    g <- ggplotGrob(plots[[1]] + theme(legend.position="bottom"))$grobs
    legend <- g[[which(sapply(g, function(x) x$name) == "guide-box")]]
    lheight <- sum(legend$height)
    plots <- lapply(plots, function(x) ggplotGrob(x + theme(legend.position = "none")))
    
    p <- gtable_cbind(gtable_rbind(plots[[1]], plots[[4]], plots[[7]]),
                      gtable_rbind(plots[[2]], plots[[5]], plots[[8]]),
                      gtable_rbind(plots[[3]], plots[[6]], plots[[9]]))
    
    grid.arrange(p, legend, 
                 nrow = 2, 
                 heights = unit.c(unit(1, "npc") - lheight, lheight))
  }

  dat_list <- vector(mode = "list", length = length(pairwise_combos))
  for(i in 1:length(pairwise_combos)){
    dat_list[[i]] <- two_way_data %>% filter(pasted == pairwise_combos[i])
  }
  
  lapply(dat_list, make_one) %>% do.call("grid_arrange_shared_legend", .)
}

fig3 <- two_way_plot(two_way_W_shredder, interesting)
grid.draw(fig3)
```
<br></br>
**Figure 3:** Heatmap showing the nine strongest interactions between pairs of parameters in the model, as determined by the GLM plotted in Figure SX. 

```{r echo=FALSE, results='hide'}
cairo_pdf(file = "figures/figure3.pdf", width = 12, height = 12); grid.draw(fig3); dev.off()
```
